### 2024-11-22 paper/paper.qmd
Uploading initial files

---
title: "Title"
subtitle: "Subtitle"
author:
  - Sameeck Bhatia
thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nullam in massa at urna ultricies lacinia eu at elit. Nam quis libero suscipit orci laoreet consequat auctor vel nunc. Curabitur viverra eleifend turpis, non mollis sem ullamcorper quis. In eros elit, tempor accumsan lectus id, lobortis interdum enim. In facilisis purus faucibus tempus pharetra. Quisque ligula dolor, cursus vehicula elit cursus, hendrerit pellentesque justo."
date: today
date-format: long
toc: true
number-sections: true
bibliography: references.bib
format: 
  pdf:
    fig.width: 6
    fig.height: 4.5
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| include: FALSE

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

Mauris ac lectus sed augue euismod blandit at non quam. Integer pharetra tortor sed purus iaculis ultrices. Sed quis sagittis urna. Suspendisse at hendrerit dui. Nunc venenatis nibh sit amet ante dictum, sollicitudin scelerisque est cursus. Etiam quis commodo sapien. Curabitur nisi mi, tempor eget sem non, faucibus lacinia metus. Quisque sed sapien ligula.

# Data

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales. Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. [@R] [@tidyverse]

## Measurement

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis. Pellentesque ut purus vitae leo rhoncus fringilla. Quisque commodo neque nec ultricies tempor.

## Outcome Variables

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

## Predictor Variables

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

# Model

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

## Model Setup

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

## Model Justification

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

# Results

## Result 1

```{r, fig.align='center'}
palmerpenguins::penguins |>
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis volutpat ut ante vestibulum sagittis. Etiam sollicitudin ligula neque, et sollicitudin erat iaculis eu. Pellentesque aliquam semper sapien, a faucibus nulla consequat vitae. Praesent in vulputate neque. Fusce fermentum tempor turpis, nec iaculis diam porta nec. Maecenas nec porta felis, at tincidunt ex. Mauris convallis tincidunt posuere. Quisque rutrum imperdiet turpis, quis bibendum est ultrices ut. Phasellus in tortor eu ligula pellentesque pellentesque. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Curabitur eleifend ac purus quis dapibus. Ut tincidunt posuere tortor, vitae venenatis nunc placerat eu. Praesent sit amet mi urna. Sed ultrices est et ligula posuere, eget molestie lorem porta. Quisque tempor, dolor porta semper aliquam, urna magna rhoncus sapien, at mattis metus nibh eleifend mi. Sed id bibendum urna, ut tempor ligula. Nulla facilisi. Nulla facilisi.

## Result 2

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = sex, y = body_mass_g, fill = sex)) +
  geom_boxplot(width = 0.5) +
  coord_flip()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis volutpat ut ante vestibulum sagittis. Etiam sollicitudin ligula neque, et sollicitudin erat iaculis eu. Pellentesque aliquam semper sapien, a faucibus nulla consequat vitae. Praesent in vulputate neque. Fusce fermentum tempor turpis, nec iaculis diam porta nec. Maecenas nec porta felis, at tincidunt ex. Mauris convallis tincidunt posuere. Quisque rutrum imperdiet turpis, quis bibendum est ultrices ut. Phasellus in tortor eu ligula pellentesque pellentesque. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Curabitur eleifend ac purus quis dapibus. Ut tincidunt posuere tortor, vitae venenatis nunc placerat eu. Praesent sit amet mi urna. Sed ultrices est et ligula posuere, eget molestie lorem porta. Quisque tempor, dolor porta semper aliquam, urna magna rhoncus sapien, at mattis metus nibh eleifend mi. Sed id bibendum urna, ut tempor ligula. Nulla facilisi. Nulla facilisi.

## Result 3

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.25) +
  coord_flip()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis volutpat ut ante vestibulum sagittis. Etiam sollicitudin ligula neque, et sollicitudin erat iaculis eu. Pellentesque aliquam semper sapien, a faucibus nulla consequat vitae. Praesent in vulputate neque. Fusce fermentum tempor turpis, nec iaculis diam porta nec. Maecenas nec porta felis, at tincidunt ex. Mauris convallis tincidunt posuere. Quisque rutrum imperdiet turpis, quis bibendum est ultrices ut. Phasellus in tortor eu ligula pellentesque pellentesque. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Curabitur eleifend ac purus quis dapibus. Ut tincidunt posuere tortor, vitae venenatis nunc placerat eu. Praesent sit amet mi urna. Sed ultrices est et ligula posuere, eget molestie lorem porta. Quisque tempor, dolor porta semper aliquam, urna magna rhoncus sapien, at mattis metus nibh eleifend mi. Sed id bibendum urna, ut tempor ligula. Nulla facilisi. Nulla facilisi. Proin mollis quam ut rutrum vulputate. Praesent commodo.

\newpage

# Discussion

## Discussion 1
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

## Discussion 2

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

## Discussion 3

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

\newpage

# Appendix {-}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc euismod nisl sollicitudin blandit fermentum. Pellentesque molestie mi augue, a consectetur tellus commodo a. Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est. Proin neque massa, interdum et laoreet vulputate, cursus vitae mi. Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor. Vivamus dictum volutpat nisl. Quisque ullamcorper erat mi, ac tempus leo efficitur non. Nulla porta id est id tempus. Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.

\newpage

# References

### 2024-11-23 scripts/simulation.R
Simulating data and running tests

@@ -1,0 +1,32 @@
+#### Preamble ####
+# Purpose: Simulates mock data on Seattle real estate sale prices.
+# Author: Sameeck Bhatia
+# Date: 22 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+
+# Setting a seed for reproducibility
+set.seed(12)
+
+n <- 100
+
+# Creating a data frame with simulated values
+simulated_data <- data.frame(
+  id = sample(10000:99999, n, replace = FALSE),
+  zip_code = sample(98101:98140, n, replace = TRUE),
+  beds = sample(c(0:5), n, replace = TRUE, prob = c(0.05, 0.1, 0.3, 0.35, 0.15, 0.05)),
+  baths = sample(c(1:4), n, replace = TRUE, prob = c(0.1, 0.2, 0.5, 0.2)),
+  sqft = round(rnorm(n, 900, 100))
+) 
+
+simulated_data <- simulated_data |>
+  mutate(price = round(sqft * rpois(n, 700), -3))
+
+# Exporting the simulated data
+write_csv(simulated_data, "data/simulated_data/simulated_data.csv")

### 2024-11-23 scripts/simulation_tests.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: Tests the validity and structure of the simulated data.
+# Author: Sameeck Bhatia
+# Date: 22 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+#   - `testthat` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+library(testthat)
+
+# Loading the simulated data
+simulated_data <- read_csv("data/simulated_data/simulated_data.csv")
+
+# Doing unit tests
+test_that("IDs are unique", {
+  expect_equal(length(unique(simulated_data$id)), n)
+})
+
+test_that("Square footage values are non-negative", {
+  expect_true(all(simulated_data$sqft > 0))
+})
+
+test_that("Square footage values are non-negative", {
+  expect_true(all(simulated_data$sqft > 0))
+})
+
+test_that("Studios have at least one bath", {
+  expect_true(any(simulated_data$beds == 0 & simulated_data$baths >= 1))
+})
+
+test_that("Number of beds and baths have at least one instance", {
+  expect_true(min(table(simulated_data$beds)) > 0)
+  expect_true(min(table(simulated_data$baths)) > 0)
+})

### 2024-11-23 scripts/cleaning.R
Loading and cleaning raw data

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Cleans the raw data to prepare for analysis.
+# Author: Sameeck Bhatia
+# Date: 23 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+#   - `janitor` package must be installed.
+#   - `arrow` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+library(janitor)
+library(arrow)
+
+# Reading the raw data
+raw_0bed <- read_csv("data/raw_data/raw_0bed.csv")
+raw_1bed <- read_csv("data/raw_data/raw_1bed.csv")
+raw_2bed <- read_csv("data/raw_data/raw_2bed.csv")
+raw_3bed <- read_csv("data/raw_data/raw_3bed.csv")
+raw_4bed <- read_csv("data/raw_data/raw_4bed.csv")
+raw_5bed <- read_csv("data/raw_data/raw_5bed.csv")
+
+# Merging the data
+raw_data <- rbind(raw_0bed, raw_1bed, raw_2bed, raw_3bed, raw_4bed, raw_5bed)
+
+# Cleaning the data
+cleaned_data <- raw_data |>
+  clean_names() |>
+  select(sale_type, mls_number, property_type, city, zip_or_postal_code, beds, 
+         baths, location, square_feet, year_built, days_on_market, hoa_month,
+         price, latitude, longitude) |>
+  filter(sale_type != "New Construction Plan",
+         !is.na(mls_number),
+         !is.na(square_feet),
+         year_built <= 2024) |>
+  rename(mls_id = mls_number,
+         zip_code = zip_or_postal_code,
+         neighbourhood = location,
+         sqft = square_feet) |>
+  mutate(half_bath = ifelse(baths - floor(baths) > 0, 1, 0),
+         baths = floor(baths),
+         property_age = 2024 - year_built,
+         price_sqft = round(price / sqft, 2))
+
+colSums(is.na(cleaned_data))
+
+cleaned_data |>
+  select(mls_id, neighbourhood, beds, baths, sqft, price) |>
+  filter(is.na(neighbourhood) | is.na(sqft))
+
+# Exporting cleaned data
+write_parquet(cleaned_data, "data/cleaned_data/cleaned_data.parquet")

### 2024-11-23 scripts/cleaning_tests.R
Updating unit tests

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Tests the validity and structure of the cleaned data.
+# Author: Sameeck Bhatia
+# Date: 23 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+#   - `testthat` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+library(testthat)
+
+# Loading the simulated data
+cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
+
+# Doing unit tests
+test_that("IDs are unique", {
+  n <- nrow(cleaned_data)
+  expect_equal(length(unique(cleaned_data$mls_id)), n)
+})
+
+test_that("ZIP code values are valid", {
+  expect_true(all(is.numeric(cleaned_data$zip_code)))
+})
+
+test_that("Number of beds and baths have at least one instance", {
+  expect_true(min(table(cleaned_data$beds)) > 0)
+  expect_true(min(table(cleaned_data$baths)) > 0)
+})
+
+test_that("Square footage values are non-negative", {
+  expect_true(all(cleaned_data$sqft > 0))
+})
+
+test_that("Year built values are valid", {
+  expect_true(all(cleaned_data$year_built <= 2024))
+})
+
+test_that("Price values are non-negative", {
+  expect_true(all(cleaned_data$price > 0))
+})
+
+test_that("Geographic coordinate values are non-zero", {
+  expect_true(all(cleaned_data$latitude != 0))
+  expect_true(all(cleaned_data$longitude != 0))
+})

### 2024-11-23 scripts/simulation_tests.R

@@ -18,6 +18,7 @@ simulated_data <- read_csv("data/simulated_data/simulated_data.csv")
 
 # Doing unit tests
 test_that("IDs are unique", {
+  n <- nrow(simulated_data)
   expect_equal(length(unique(simulated_data$id)), n)
 })
 

### 2024-11-23 scripts/cleaning.R
Updating scripts and data

@@ -33,7 +33,6 @@ cleaned_data <- raw_data |>
          baths, location, square_feet, year_built, days_on_market, hoa_month,
          price, latitude, longitude) |>
   filter(sale_type != "New Construction Plan",
-         !is.na(mls_number),
          !is.na(square_feet),
          year_built <= 2024) |>
   rename(mls_id = mls_number,
@@ -43,7 +42,8 @@ cleaned_data <- raw_data |>
   mutate(half_bath = ifelse(baths - floor(baths) > 0, 1, 0),
          baths = floor(baths),
          property_age = 2024 - year_built,
-         price_sqft = round(price / sqft, 2))
+         price_sqft = round(price / sqft, 2),
+         property_type = recode(property_type, "Single Family Residential" = "Single Family"))
 
 colSums(is.na(cleaned_data))
 

### 2024-11-23 scripts/cleaning_tests.R

@@ -13,7 +13,7 @@
 library(tidyverse)
 library(testthat)
 
-# Loading the simulated data
+# Reading the cleaned data
 cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
 
 # Doing unit tests

### 2024-11-23 scripts/simulation_tests.R

@@ -13,7 +13,7 @@
 library(tidyverse)
 library(testthat)
 
-# Loading the simulated data
+# Reading the simulated data
 simulated_data <- read_csv("data/simulated_data/simulated_data.csv")
 
 # Doing unit tests

### 2024-11-23 scripts/eda.R
Completing EDA

@@ -1,0 +1,92 @@
+#### Preamble ####
+# Purpose: Conducting exploratory data analysis on the cleaned data.
+# Author: Sameeck Bhatia
+# Date: 23 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+library(arrow)
+
+# Reading the cleaned data
+cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
+
+glimpse(cleaned_data)
+
+# Displaying summary statistics
+summary(cleaned_data)
+
+# Plotting histograms
+ggplot(cleaned_data, aes(x = sqft)) + 
+  geom_histogram(fill = "beige", color = "black") +
+  labs(x = "Size (sqft)", y = NULL)
+
+ggplot(cleaned_data, aes(x = days_on_market)) + 
+  geom_histogram(fill = "beige", color = "black") +
+  labs(x = "Days on Market", y = NULL)
+
+ggplot(cleaned_data, aes(x = hoa_month)) + 
+  geom_histogram(fill = "beige", color = "black") +
+  labs(x = "HOA Fees ($/month)", y = NULL)
+
+ggplot(cleaned_data, aes(x = price)) + 
+  geom_histogram(fill = "beige", color = "black") +
+  labs(x = "Price ($)", y = NULL)
+
+ggplot(cleaned_data, aes(x = property_age)) + 
+  geom_histogram(fill = "beige", color = "black") +
+  labs(x = "Property Age (years)", y = NULL)
+
+# Plotting bar charts
+cleaned_data |> 
+  ggplot(aes(x = property_type)) +
+  geom_bar(fill = "beige", color = "black") +
+  labs(x = "Property Type", y = NULL)
+
+cleaned_data |> 
+  mutate(beds = as.factor(beds)) |>
+  ggplot(aes(x = beds)) +
+  geom_bar(fill = "beige", color = "black") +
+  labs(x = "Beds", y = NULL)
+
+cleaned_data |> 
+  mutate(baths = as.factor(baths)) |>
+  ggplot(aes(x = baths)) +
+  geom_bar(fill = "beige", color = "black") +
+  labs(x = "Baths", y = NULL)
+
+# Plotting box plots
+cleaned_data |> 
[…]

### 2024-11-24 scripts/eda.R
Updating comments

@@ -6,6 +6,7 @@
 # License: MIT
 # Pre-requisites:
 #   - `tidyverse` package must be installed.
+#   - `arrow` package must be installed.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-24 scripts/modeling.R
Fitting and evaluating models

@@ -1,0 +1,88 @@
+#### Preamble ####
+# Purpose: Preparing and modeling the clean data for training and testing.
+# Author: Sameeck Bhatia
+# Date: 23 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   - `tidyverse` package must be installed.
+#   - `arrow` package must be installed.
+#   - `rsample` package must be installed.
+#   - `MLmetrics` package must be installed.
+#   - `car` package must be installed.
+# Any other information needed? None
+
+# Loading required libraries
+library(tidyverse)
+library(arrow)
+library(rsample)
+library(MLmetrics)
+library(car)
+
+# Reading the cleaned data
+cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
+
+# Preparing data for the model
+model_data <- cleaned_data |>
+  select(property_type, zip_code, beds, baths, half_bath, sqft, days_on_market,
+         property_age, price)
+
+set.seed(12)
+
+split <- initial_split(model_data, prop = 0.7)
+
+training_data <- training(split)
+testing_data <- testing(split)
+
+# Fitting model 1
+model1 <- lm(data = training_data, 
+             formula = price ~ .)
+
+y_pred <- predict(model1, newdata = testing_data)
+y_true <- testing_data$price
+
+rmse1 <- RMSE(y_pred, y_true)
+mae1 <- MAE(y_pred, y_true)
+r2_1 <- R2_Score(y_pred, y_true)
+
+summary(model1)
+
+plot(fitted(model1), resid(model1))
+qqnorm(resid(model1));qqline(resid(model1))
+powerTransform(model1)
+
+# Fitting model 2
+model2 <- lm(data = training_data, 
+             formula = price ~ property_type + beds + sqft + property_age)
+
+y_pred <- predict(model2, newdata = testing_data)
+y_true <- testing_data$price
+
+rmse2 <- RMSE(y_pred, y_true)
+mae2 <- MAE(y_pred, y_true)
+r2_2 <- R2_Score(y_pred, y_true)
+
+summary(model2)
+
+anova(model1, model2)
+
+# Fitting model 3
+model3 <- lm(data = training_data, 
[…]

### 2024-11-25 scripts/modeling.R
Exporting the model

@@ -27,6 +27,7 @@ model_data <- cleaned_data |>
   select(property_type, zip_code, beds, baths, half_bath, sqft, days_on_market,
          property_age, price)
 
+# Setting seed for reproducibility
 set.seed(12)
 
 split <- initial_split(model_data, prop = 0.7)
@@ -41,15 +42,12 @@ model1 <- lm(data = training_data,
 y_pred <- predict(model1, newdata = testing_data)
 y_true <- testing_data$price
 
+# Evaluating model 1
 rmse1 <- RMSE(y_pred, y_true)
 mae1 <- MAE(y_pred, y_true)
 r2_1 <- R2_Score(y_pred, y_true)
 
 summary(model1)
-
-plot(fitted(model1), resid(model1))
-qqnorm(resid(model1));qqline(resid(model1))
-powerTransform(model1)
 
 # Fitting model 2
 model2 <- lm(data = training_data, 
@@ -58,6 +56,7 @@ model2 <- lm(data = training_data,
 y_pred <- predict(model2, newdata = testing_data)
 y_true <- testing_data$price
 
+# Evaluating model 2
 rmse2 <- RMSE(y_pred, y_true)
 mae2 <- MAE(y_pred, y_true)
 r2_2 <- R2_Score(y_pred, y_true)
@@ -66,6 +65,10 @@ summary(model2)
 
 anova(model1, model2)
 
+plot(fitted(model2), resid(model2))
+qqnorm(resid(model2));qqline(resid(model2))
+powerTransform(model2)
+
 # Fitting model 3
 model3 <- lm(data = training_data, 
              formula = price^(-1/4) ~ property_type + beds + sqft + property_age + sqft:beds)
@@ -73,11 +76,15 @@ model3 <- lm(data = training_data,
 y_pred <- predict(model3, newdata = testing_data)
 y_true <- testing_data$price^(-1/4)
 
+# Evaluating model 3
 rmse3 <- RMSE(y_pred, y_true)
 mae3 <- MAE(y_pred, y_true)
 r2_3 <- R2_Score(y_pred, y_true)
 
 summary(model3)
+
+plot(fitted(model3), resid(model3))
+qqnorm(resid(model3));qqline(resid(model3))
 
 # Summarizing metrics
 data.frame(
@@ -86,3 +93,6 @@ data.frame(
   model3 = c(rmse3, mae3, r2_3),
   row.names = c("RMSE", "MAE", "R^2")
 )
+
+# Exporting model
+saveRDS(model3, "models/valuation_model.rds")

### 2024-11-25 paper/paper.qmd
Updating references, title, and data section

@@
 ---
-title: "Title"
+title: "Valuing Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach"
 subtitle: "Subtitle"
@@
   - Sameeck Bhatia
+date: today
+date-format: long
 thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
@@
 Quisque ligula dolor, cursus vehicula elit cursus, hendrerit pellentesque justo."
-date: today
-date-format: long
 toc: true
@@
 library(palmerpenguins)
+library(kableExtra)
+library(arrow)
+library(stargazer)
+library(ggpubr)
 ```
@@
 Suspendisse lacinia ornare consectetur.
-Quisque ultricies tortor non ullamcorper porta.
-Mauris ac leo quis lacus tincidunt cursus ut id felis.
 
-Mauris ac lectus sed augue euismod blandit at non quam.
-Integer pharetra tortor sed purus iaculis ultrices.
-Sed quis sagittis urna.
-Suspendisse at hendrerit dui.
-Nunc venenatis nibh sit amet ante dictum, sollicitudin scelerisque est cursus.
-Etiam quis commodo sapien.
-Curabitur nisi mi, tempor eget sem non, faucibus lacinia metus.
-Quisque sed sapien ligula.
+Donec commodo diam non arcu facilisis, sit amet egestas massa consequat.
+Vestibulum semper congue libero, eu vestibulum mi cursus sed.
+Morbi ut justo quis ligula molestie mattis.
+In luctus, dolor ac posuere pretium, ex ex sollicitudin ligula, a varius libero nisi sed arcu.
+Nulla et bibendum nunc.
+Morbi sit amet lorem sit amet lectus bibendum tempus aliquam sit amet lectus.
+Sed ligula nulla, pulvinar quis ipsum a, condimentum egestas mi.
+Aenean libero tortor, vehicula vitae velit sit amet, eleifend semper elit.
+Nam non fringilla nisi.
+Vivamus quis consequat odio, et consectetur risus.
+Proin vulputate tristique mauris vitae interdum.
+Phasellus efficitur dictum nibh, id dictum nisl rutrum et.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
-Curabitur ac leo non nunc ultrices dignissim.
-Phasellus faucibus elit id nulla pretium volutpat.
-Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo.
-Curabitur vulputate imperdiet nulla eu finibus.
-Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam.
-Etiam nunc libero, faucibus nec porta a, dignissim a nisi.
-Phasellus laoreet elementum turpis a vestibulum.
-Quisque in urna varius, lobortis eros eu, blandit sapien.
-In lobortis orci id nunc facilisis, non tempor nisl ultricies.
-Sed pellentesque tortor cursus ipsum fringilla interdum.
-Cras lacinia tempor placerat.
-Proin eu elit elit.
-Sed molestie dolor et posuere sodales.
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-[@R] [@tidyverse]
+## Overview
+
+The data used in this paper comes from Redfin, a real estate brokerage and mortgage company, and represents current property prices across Seattle.
+While the data was collected from @redfin, similar data could have been sourced from competitors like Zillow and Realtor.com, with Zillow being the largest of the three.
+If data from Zillow or Realtor.com had been used, the listings might vary slightly since each platform likely features different sellers and listings.
+However, Redfin was chosen as the source because it is the only brokerage firm that allows public downloads of its listings, helping to avoid violations of data extraction policies.
+
+The dataset includes only properties within Seattle’s official city boundaries and covers condos, townhouses, and single-family homes currently on the market.
+The raw data contains 27 variables, 18 of which have been included in the cleaned dataset.
+These include variables such as MLS number, number of bedrooms, neighborhood, and geographic coordinates.
+Additionally, some variables were constructed specifically for this analysis, such as `half_bath`, `property_age`, and `price_sqft`.
+The first two were added to enhance the dataset and valuation model, while `price_sqft` was created as an alternate response variable.
+Further details about these variables can be found in Appendix A.
+The data was analyzed using `R` [@R] and the `tidyverse` [@tidyverse] package, while visualizations have been created using `tidyverse`.
+
+## Summary Statistics
+
+```{r}
+#| output: asis
+
+data <- read_parquet("../data/cleaned_data/cleaned_data.parquet") |>
+  as.data.frame()
+
+data |>
+  select(!c("zip_code", "latitude", "longitude")) |>
+  stargazer(type = "latex", header = FALSE, digits = 2, median = TRUE, 
+            label = "tbl-stats")
+```
+
+Table \ref{tbl-stats} presents summary statistics for all original and derived numeric variables in the dataset.
+The mean number of bedrooms is 2.69, with a median of 3, while the mean number of bathrooms is 2.03, with a median of 2.
+This indicates that the typical listing has around three bedrooms and two bathrooms, commonly seen in townhouses and single-family homes.
+The average property size is approximately 1,710 square feet, with a median of 1,373 square feet, suggesting a positive skew in property size due to a few larger homes in the dataset.
+The average time a property remains on the market is around 64 days (just over 2 months), with a maximum of 878 days (nearly 2.5 years), indicating low demand in Seattle’s real estate market, especially since the data contains only active listings.
+The mean and median property prices are \$1,107,431 and \$775,000, respectively, while the mean and median price per square foot are \$631.41 and \$590.71, respectively.
+These figures highlight Seattle as one of the most expensive residential markets in the United States.
 
@@
 
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-Donec at malesuada purus.
-Cras ultrices vitae metus eget aliquet.
-Aliquam sollicitudin elit placerat, porta mauris in, semper felis.
-Nulla vestibulum dolor sed tempor efficitur.
-In bibendum libero ex, in sollicitudin arcu maximus porttitor.
-Pellentesque ac sapien sit amet magna ornare auctor.
-In hac habitasse platea dictumst.
-Cras ut urna dapibus elit mattis aliquam non sed augue.
-Suspendisse lacinia ornare consectetur.
-Quisque ultricies tortor non ullamcorper porta.
-Mauris ac leo quis lacus tincidunt cursus ut id felis.
-Pellentesque ut purus vitae leo rhoncus fringilla.
-Quisque commodo neque nec ultricies tempor.
+In the United States, buyers and sellers have the freedom to select the real estate agent or brokerage firm they wish to work with for transactions.
+Consequently, agents often represent multiple listings within their region or city.
+Agents receive detailed information on each property from real estate appraisers, who measure variables such as the number of bedrooms, bathrooms, and square footage [@nar].
+These measurement practices, except for price, are strictly regulated to ensure accuracy for all stakeholders.
+
+For property prices, appraisers typically estimate values based on the prices of recently sold comparable properties and the specific characteristics of the property.
+This valuation process is less regulated, as it serves primarily as a reference point for buyers and sellers.
+Real estate agents may gather price estimates from multiple appraisers to calculate an average.
+The prices observed in the data, although guided by these values, are ultimately set by the seller.
+All this information is uploaded to the Multiple Listing Service (MLS), a private database accessible only to agents and brokerage firms via subscription fees [@bankrate].
+However, U.S.
+laws allow companies like Redfin, Trulia, and Zillow to extract and share MLS data with the public, fostering competition and transparency.
+For this study, data was sourced from the Seattle region’s MLS database.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-out ipsum dolor sit amet, consectetur adipiscing elit.
 Nunc euismod nisl sollicitudin blandit fermentum.
@@
 
+```{r fig-out, fig.width=7, fig.height=2.8, fig.cap='Distributions of Property Price and Logarithmic Property Price'}
+p1 <- data |>
+  select(price) |>
+  ggplot(aes(x = price)) +
+  geom_histogram(binwidth = 1000000, fill = "beige", color = "black") +
+  labs(x = "Price ($)", y = "Count")
+
+p2 <- data |>
+  select(price) |>
+  ggplot(aes(x = log(price))) +
+  geom_histogram(binwidth = 0.25, fill = "beige", color = "black") +
+  labs(x = " ln(Price) ($)", y = "Count")
+
+ggarrange(p1, p2, labels = c("A", "B"))
+```
+
 ## Predictor Variables
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-pred1 ipsum dolor sit amet, consectetur adipiscing elit.
 Nunc euismod nisl sollicitudin blandit fermentum.
@@
 Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+
+```{r fig-pred1, fig.width=7, fig.height=2.9, fig.cap='Distributions of Property Type and Number of Bedrooms'}
+p3 <- data |> 
+  ggplot(aes(x = property_type)) +
+  geom_bar(fill = "beige", color = "black") +
+  labs(x = "Property Type", y = "Count")
+
+p4 <- data |> 
+  mutate(beds = as.factor(beds)) |>
+  ggplot(aes(x = beds)) +
+  geom_bar(fill = "beige", color = "black") +
+  labs(x = "Beds", y = "Count")
+
+ggarrange(p3, p4, labels = c("C", "D"))
+```
+
+@fig-pred2 ipsum dolor sit amet, consectetur adipiscing elit.
+Nunc euismod nisl sollicitudin blandit fermentum.
+Pellentesque molestie mi augue, a consectetur tellus commodo a.
+Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
+Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
+Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
+Vivamus dictum volutpat nisl.
+Quisque ullamcorper erat mi, ac tempus leo efficitur non.
+Nulla porta id est id tempus.
+Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+
+```{r fig-pred2, fig.width=7, fig.height=2.8, fig.cap='Distributions of Property Size and Property Age'}
+p5 <- data |>
+  ggplot(aes(x = sqft)) + 
+  geom_histogram(binwidth = 500, fill = "beige", color = "black") +
+  labs(x = "Size (sqft)", y = NULL)
+
+p6 <- data |> 
+  ggplot(aes(x = property_age)) + 
+  geom_histogram(binwidth = 10, fill = "beige", color = "black") +
+  labs(x = "Property Age (years)", y = NULL)
+
+ggarrange(p5, p6, labels = c("E", "F"))
+```
 
@@
 Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
 
@@
 
+\appendix
+
 # Appendix {-}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+# Additional Data Details
+
+```{r}
+data_dictionary <- data.frame(
+  Variable = c(
+    "SALE TYPE", "SOLD DATE", "PROPERTY TYPE", "ADDRESS", "CITY", 
+    "STATE OR PROVINCE", "ZIP OR POSTAL CODE", "PRICE", "BEDS", 
+    "BATHS", "LOCATION", "SQUARE FEET", "LOT SIZE", "YEAR BUILT", 
+    "DAYS ON MARKET", "$/SQUARE FEET", "HOA/MONTH", "STATUS", 
+    "NEXT OPEN HOUSE START TIME", "NEXT OPEN HOUSE END TIME", 
+    "URL", "SOURCE", "MLS#", "FAVORITE", "INTERESTED", "LATITUDE", "LONGITUDE"
+  ),
+  Description = c(
+    "Type of sale (e.g., MLS listing, new construction home, new construction plan).",
+    "Date when the property was sold.",
+    "Type of property (e.g., condo, single-family house, townhouse).",
+    "Full address of the property.",
+    "City where the property is located.",
+    "State or province where the property is located.",
+    "ZIP code of the property location.",
+    "Sale price of the property.",
+    "Number of bedrooms in the property.",
+    "Number of bathrooms in the property.",
+    "Neighbourhood of the property.",
+    "Total square footage of the property.",
+    "Lot size in square feet",
+    "Year the property was built.",
+    "Number of days the property has been listed on the market.",
+    "Price per square foot of the property.",
+    "Monthly Homeowners Association (HOA) fee, if applicable.",
+    "Current status of the property (e.g., sold, pending, active).",
+    "Start time of the next scheduled open house, if available.",
+    "End time of the next scheduled open house, if available.",
+    "URL to additional property details.",
+    "Source of the property data (e.g., MLS, Zillow).",
+    "Multiple Listing Service (MLS) identification number for the property.",
+    "Indicates whether the property is marked as a favorite (e.g., Y/N).",
+    "Indicates whether the user has expressed interest in the property (e.g., Y/N).",
+    "Latitude of the property location for geospatial analysis.",
+    "Longitude of the property location for geospatial analysis."
+  )
+)
+
+kable(data_dictionary, col.names = c("Variable", "Description"), align = "l")
+```
 

### 2024-11-25 scripts/simulation.R
Updating simulation data and tests

@@ -22,6 +22,8 @@ simulated_data <- data.frame(
   zip_code = sample(98101:98140, n, replace = TRUE),
   beds = sample(c(0:5), n, replace = TRUE, prob = c(0.05, 0.1, 0.3, 0.35, 0.15, 0.05)),
   baths = sample(c(1:4), n, replace = TRUE, prob = c(0.1, 0.2, 0.5, 0.2)),
+  property_age = round(rexp(n, 1/5)),
+  days_on_market = round(rexp(n, 1/14)),
   sqft = round(rnorm(n, 900, 100))
 ) 
 

### 2024-11-25 scripts/simulation_tests.R

@@ -22,12 +22,20 @@ test_that("IDs are unique", {
   expect_equal(length(unique(simulated_data$id)), n)
 })
 
-test_that("Square footage values are non-negative", {
+test_that("Square footage values are positive", {
   expect_true(all(simulated_data$sqft > 0))
 })
 
-test_that("Square footage values are non-negative", {
+test_that("Square footage values are positive", {
   expect_true(all(simulated_data$sqft > 0))
+})
+
+test_that("Age values are non-negative", {
+  expect_true(all(simulated_data$property_age >= 0))
+})
+
+test_that("Days on market values are non-negative", {
+  expect_true(all(simulated_data$days_on_market >= 0))
 })
 
 test_that("Studios have at least one bath", {

### 2024-11-26 paper/paper.qmd
Completing data section

@@
 abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nullam in massa at urna ultricies lacinia eu at elit.
-Nam quis libero suscipit orci laoreet consequat auctor vel nunc.
-Curabitur viverra eleifend turpis, non mollis sem ullamcorper quis.
-In eros elit, tempor accumsan lectus id, lobortis interdum enim.
-In facilisis purus faucibus tempus pharetra.
-Quisque ligula dolor, cursus vehicula elit cursus, hendrerit pellentesque justo."
+Morbi faucibus ipsum dolor, in tincidunt justo tempor sit amet proin non turpis a orci imperdiet pretium.
+Aenean aliquam eros eu consectetur cursus.
+Proin vel tempor ante vestibulum felis nisi, ultricies sed porttitor in, viverra nec augue.
+Sed convallis a tortor in pretium.
+Vivamus metus ipsum, faucibus vitae turpis sed, sodales auctor dolor sed a orci id orci pulvinar tincidunt sed ut tellus.
+Morbi venenatis tempor arcu, a auctor dolor congue sit amet.
+Nam auctor tincidunt ligula rutrum aliquam.
+Sed facilisis at nulla et posuere."
 toc: true
@@
 In lobortis orci id nunc facilisis, non tempor nisl ultricies.
-Sed pellentesque tortor cursus ipsum fringilla interdum.
-Cras lacinia tempor placerat.
-Proin eu elit elit.
-Sed molestie dolor et posuere sodales.
 
@@
 laws allow companies like Redfin, Trulia, and Zillow to extract and share MLS data with the public, fostering competition and transparency.
-For this study, data was sourced from the Seattle region’s MLS database.
 
@@
 
-@fig-out ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+The primary goal of this paper is to estimate the outcome variable, price, using a valuation model.
+This variable represents the market value of active listings at the time of data collection, focusing exclusively on properties located in Seattle.
+@fig-out provides insights into the distribution of property prices.
+Plot A shows that property prices are highly skewed, with a maximum value near \$40,000,000 and several listings exceeding \$5,000,000.
+To better understand the distribution, Plot B presents the logarithmic transformation of property prices.
+While the transformed distribution is less extreme, it remains skewed, indicating the presence of properties with exceptionally high valuations in the Seattle real estate market.
 
-```{r fig-out, fig.width=7, fig.height=2.8, fig.cap='Distributions of Property Price and Logarithmic Property Price'}
+```{r fig-out, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Price and Logarithmic Property Price'}
 p1 <- data |>
@@
 
-@fig-pred1 ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+**Property Type (`property_type`):** This categorical variable identifies the type of property for each observation, based on classifications set by the MLS.
+The three main property types included in the analysis are "Condo/Co-op," "Single Family," and "Townhouse." These classifications are critical for valuation, as different property types possess distinct features that influence their market value.
 
-```{r fig-pred1, fig.width=7, fig.height=2.9, fig.cap='Distributions of Property Type and Number of Bedrooms'}
+**Number of Bedrooms (`beds`):** This numeric variable represents the count of full bedrooms in a property, as measured by appraisers.
+It is a key factor in the analysis, as more bedrooms often correlate with greater living space and the potential to accommodate larger households.
+
+**Property Size (`sqft`):** Measured in square feet, this variable reflects the total size of the property and is determined by appraisers.
+As the United States predominantly uses the imperial system, square footage is a standard unit.
+Larger properties generally hold more value, making this a significant predictor in property valuation.
+
+**Property Age (`property_age`):** This derived variable calculates the age of a property in years, based on the difference between the year of data collection and the `year_built`.
+Age is a vital consideration in determining price, as newer properties are often smaller due to rising construction costs and increasingly strict zoning regulations.
+
+@fig-pred1 represents the distributions of property types and number of bedrooms.
+Plot C shows that single-family homes are the most common, followed closely by condos, with townhouses being the least prevalent.
+This likely reflects the mix of apartments and houses present in Seattle's real estate market.
+Plot D highlights that properties with 2 or 3 bedrooms are most frequent, while some properties even have north of 7 bedrooms.
+This is likely because condos typically feature 1-2 bedrooms, whereas single-family homes and townhouses commonly have 3-4 bedrooms.
+
+```{r fig-pred1, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Type and Number of Bedrooms'}
 p3 <- data |> 
@@
 
-@fig-pred2 ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+@fig-pred2 displays the distributions of property size and age.
+Plot E shows that property size is positively skewed, with most properties under 2,500 square feet and a peak between 700 and 1,200 square feet.
+Properties exceeding 5,000 square feet likely represent luxury homes or mansions.
+Plot F indicates that property age is also positively skewed, with the most common properties being less than 10 years old.
+The second most common group is 20-30 years old, likely reflecting construction surges before the 2008 financial crisis and real estate market downturn.
 
-```{r fig-pred2, fig.width=7, fig.height=2.8, fig.cap='Distributions of Property Size and Property Age'}
+```{r fig-pred2, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Size and Property Age'}
 p5 <- data |>

### 2024-11-26 scripts/cleaning.R
Updating data, scripts, and models

@@ -33,17 +33,19 @@ cleaned_data <- raw_data |>
          baths, location, square_feet, year_built, days_on_market, hoa_month,
          price, latitude, longitude) |>
   filter(sale_type != "New Construction Plan",
-         !is.na(square_feet),
+         !is.na(square_feet), !is.na(year_built),
          year_built <= 2024) |>
   rename(mls_id = mls_number,
          zip_code = zip_or_postal_code,
          neighbourhood = location,
          sqft = square_feet) |>
-  mutate(half_bath = ifelse(baths - floor(baths) > 0, 1, 0),
+  mutate(hoa_month = replace_na(hoa_month, 0),
+         half_bath = ifelse(is.numeric(baths) & (baths - floor(baths) > 0), 1, 0),
          baths = floor(baths),
          property_age = 2024 - year_built,
          price_sqft = round(price / sqft, 2),
-         property_type = recode(property_type, "Single Family Residential" = "Single Family"))
+         property_type = case_when(property_type == "Single Family Residential" ~ "Single Family",
+                                   .default = property_type))
 
 colSums(is.na(cleaned_data))
 

### 2024-11-26 scripts/modeling.R

@@ -9,7 +9,6 @@
 #   - `arrow` package must be installed.
 #   - `rsample` package must be installed.
 #   - `MLmetrics` package must be installed.
-#   - `car` package must be installed.
 # Any other information needed? None
 
 # Loading required libraries
@@ -17,7 +16,6 @@ library(tidyverse)
 library(arrow)
 library(rsample)
 library(MLmetrics)
-library(car)
 
 # Reading the cleaned data
 cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
@@ -25,7 +23,7 @@ cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
 # Preparing data for the model
 model_data <- cleaned_data |>
   select(property_type, zip_code, beds, baths, half_bath, sqft, days_on_market,
-         property_age, price)
+         hoa_month, property_age, price)
 
 # Setting seed for reproducibility
 set.seed(12)
@@ -39,27 +37,11 @@ testing_data <- testing(split)
 model1 <- lm(data = training_data, 
              formula = price ~ .)
 
-y_pred <- predict(model1, newdata = testing_data)
-y_true <- testing_data$price
-
-# Evaluating model 1
-rmse1 <- RMSE(y_pred, y_true)
-mae1 <- MAE(y_pred, y_true)
-r2_1 <- R2_Score(y_pred, y_true)
-
 summary(model1)
 
 # Fitting model 2
 model2 <- lm(data = training_data, 
-             formula = price ~ property_type + beds + sqft + property_age)
-
-y_pred <- predict(model2, newdata = testing_data)
-y_true <- testing_data$price
-
-# Evaluating model 2
-rmse2 <- RMSE(y_pred, y_true)
-mae2 <- MAE(y_pred, y_true)
-r2_2 <- R2_Score(y_pred, y_true)
+             formula = price ~ property_type + beds + sqft + property_age + hoa_month)
 
 summary(model2)
 
@@ -70,29 +52,28 @@ qqnorm(resid(model2));qqline(resid(model2))
 powerTransform(model2)
 
 # Fitting model 3
-model3 <- lm(data = training_data, 
-             formula = price^(-1/4) ~ property_type + beds + sqft + property_age + sqft:beds)
-
-y_pred <- predict(model3, newdata = testing_data)
-y_true <- testing_data$price^(-1/4)
-
[…]

### 2024-11-27 scripts/modeling.R
Renaming models

@@ -22,7 +22,7 @@ cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
 
 # Preparing data for the model
 model_data <- cleaned_data |>
-  select(property_type, zip_code, beds, baths, half_bath, sqft, days_on_market,
+  select(property_type, beds, baths, half_bath, sqft, days_on_market,
          hoa_month, property_age, price)
 
 # Setting seed for reproducibility
@@ -49,7 +49,6 @@ anova(model1, model2)
 
 plot(fitted(model2), resid(model2))
 qqnorm(resid(model2));qqline(resid(model2))
-powerTransform(model2)
 
 # Fitting model 3
 model3 <- lm(data = training_data, formula = price ~ beds + sqft + hoa_month + sqft:beds)
@@ -75,5 +74,5 @@ R2_Score(testing_data$pred3, testing_data$price)
 
 # Exporting models
 saveRDS(model1, "models/full_model.rds")
-saveRDS(model2, "models/partial_model.rds")
+saveRDS(model2, "models/reduced_model.rds")
 saveRDS(model3, "models/final_model.rds")

### 2024-11-27 paper/paper.qmd
Updating introduction and model sections

@@
 library(ggpubr)
+library(modelsummary)
 ```
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
-Curabitur ac leo non nunc ultrices dignissim.
-Phasellus faucibus elit id nulla pretium volutpat.
-Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo.
-Curabitur vulputate imperdiet nulla eu finibus.
-Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam.
-Etiam nunc libero, faucibus nec porta a, dignissim a nisi.
-Phasellus laoreet elementum turpis a vestibulum.
-Quisque in urna varius, lobortis eros eu, blandit sapien.
-In lobortis orci id nunc facilisis, non tempor nisl ultricies.
+Seattle’s real estate market has experienced significant fluctuations since the start of the millennium.
+Despite these booms and busts, property prices in Seattle remain among the highest in the United States.
+Demand has steadily increased, and even during periods of lower demand, the market continues to attract buyers.
+At current price levels, more potential homebuyers are seeking reliable information to find the best deals.
+With advancements in technology and the increasing availability of data, tools for assessing property values have become essential.
+One critical tool is an estimate of a property’s fair value, which can help buyers determine whether they are overpaying or underpaying relative to the market.
 
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-Donec at malesuada purus.
-Cras ultrices vitae metus eget aliquet.
-Aliquam sollicitudin elit placerat, porta mauris in, semper felis.
-Nulla vestibulum dolor sed tempor efficitur.
-In bibendum libero ex, in sollicitudin arcu maximus porttitor.
-Pellentesque ac sapien sit amet magna ornare auctor.
-In hac habitasse platea dictumst.
-Cras ut urna dapibus elit mattis aliquam non sed augue.
-Suspendisse lacinia ornare consectetur.
+This paper focuses on developing an accurate and accessible property valuation model tailored to Seattle’s housing market.
+The model incorporates current market data to provide fair value estimates for individual properties.
+The goal is to inform readers about the key factors driving real estate prices in Seattle while empowering them with a tool to make better home-buying decisions.
+By applying machine learning and statistical techniques, the model aims to ensure accuracy and reflect the intrinsic value of properties rather than merely their market prices.
 
-Donec commodo diam non arcu facilisis, sit amet egestas massa consequat.
-Vestibulum semper congue libero, eu vestibulum mi cursus sed.
-Morbi ut justo quis ligula molestie mattis.
-In luctus, dolor ac posuere pretium, ex ex sollicitudin ligula, a varius libero nisi sed arcu.
-Nulla et bibendum nunc.
-Morbi sit amet lorem sit amet lectus bibendum tempus aliquam sit amet lectus.
-Sed ligula nulla, pulvinar quis ipsum a, condimentum egestas mi.
-Aenean libero tortor, vehicula vitae velit sit amet, eleifend semper elit.
-Nam non fringilla nisi.
-Vivamus quis consequat odio, et consectetur risus.
-Proin vulputate tristique mauris vitae interdum.
-Phasellus efficitur dictum nibh, id dictum nisl rutrum et.
+Existing valuation tools, such as Zillow’s “Zestimate” and Redfin’s proprietary estimates, have limitations.
+These models are often closed-source, leaving potential users unable to access the methodology or even the results without cost.
+This creates a significant gap in the availability of open, free, and transparent valuation tools.
+The absence of such tools denies buyers the advantage of comprehensive, unbiased information to assess property values independently.
+To address this gap, data was collected from Redfin’s semi-public dataset on current Seattle property listings.
+The analysis was conducted using the R programming language for data cleaning, testing, model creation, and result interpretation.
 
-# Data
+**TO-DO: ** Preliminary findings, significance, estimand
+
+The paper is structured as follows: @sec-data introduces the dataset and variables.
+@sec-model outlines the model design and its significance.
+@sec-results presents the model’s predictions and actionable insights for buyers.
+Finally, @sec-discuss explores potential applications of the model in other cities, factors influencing property prices, and the generalizability of the model.
+
+# Data {#sec-data}
 
@@
 
-```{r fig-out, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Price and Logarithmic Property Price'}
+```{r fig-out, fig.width=7, fig.height=2.5, fig.cap='Heavily Skewed Property Prices and Skewed Still for Logarithmic Prices'}
 p1 <- data |>
@@
   geom_histogram(binwidth = 0.25, fill = "beige", color = "black") +
-  labs(x = " ln(Price) ($)", y = "Count")
+  labs(x = " ln(Price)", y = "Count")
 
@@
 
-```{r fig-pred1, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Type and Number of Bedrooms'}
+```{r fig-pred1, fig.width=7, fig.height=2.5, fig.cap='Single Family Homes and 2 or 3 Bedrooms Properties Most Listing on the Market'}
 p3 <- data |> 
@@
 
-```{r fig-pred2, fig.width=7, fig.height=2.6, fig.cap='Distributions of Property Size and Property Age'}
+```{r fig-pred2, fig.width=7, fig.height=2.5, fig.cap='Skewed Distribution of Sizes and Most Properties Built Over the Last Decade'}
 p5 <- data |>
@@
 
-# Model
+\newpage
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+# Model {#sec-model}
 
-## Model Setup
+## Overview
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
+The model uses a multiple linear regression approach to estimate property values in Seattle.
+It incorporates features like the number of bedrooms (`beds`), square footage (`sqft`), homeowner association fees (`hoa_month`), and an interaction term (`sqft` $\times$ `beds`) to predict prices.
+These characteristics were selected for their significance in influencing property valuations.
+The final model was chosen after thorough validation, achieving a good balance of accuracy and interpretability.
 
-## Model Justification
+## Setup
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc euismod nisl sollicitudin blandit fermentum.
-Pellentesque molestie mi augue, a consectetur tellus commodo a.
-Sed dui lectus, sollicitudin ut sagittis eget, venenatis eget est.
-Proin neque massa, interdum et laoreet vulputate, cursus vitae mi.
-Ut scelerisque, dui vitae consequat fringilla, nunc felis pharetra dui, ac condimentum nibh ipsum non tortor.
-Vivamus dictum volutpat nisl.
-Quisque ullamcorper erat mi, ac tempus leo efficitur non.
-Nulla porta id est id tempus.
-Quisque pulvinar, dolor eget elementum tincidunt, est neque bibendum purus, quis sodales ipsum eros ac turpis.
+The linear regression model can be represented as:
 
-# Results
+$$
+\text{price} = \beta_0 + \beta_1 \cdot \text{beds} + \beta_2 \cdot \text{sqft} + \beta_3 \cdot \text{hoa\_month} + \beta_4 \cdot (\text{sqft} \cdot \text{beds})
+$$
+
+, where the coefficients are described as:
+
+**$\beta_0$: Intercept**
+
+The predicted value of `price` when all predictors ($\text{beds}$, $\text{sqft}$, $\text{hoa\_month}$, $\text{sqft} \cdot \text{beds}$) are zero.
+
+**$\beta_1$: Coefficient for `beds`**
+
+The effect of adding one more bedroom ($\text{beds}$) on `price`, holding other factors constant (except for the interaction term).
+
+**$\beta_2$: Coefficient for `sqft`**
+
+The effect of increasing square footage ($\text{sqft}$) by one unit on `price`, holding other factors constant (except for the interaction term).
+
+**$\beta_3$: Coefficient for `hoa_month`**
+
+The effect of a one-unit increase in monthly homeowner's association fees (`hoa_month`) on `price`, holding other factors constant.
+
+**$\beta_4$: Interaction between `sqft` and `beds`**
+
+Represents how the relationship between $\text{sqft}$ and `price` changes depending on the number of bedrooms ($\text{beds}$).
+
+## Limitations
+
+This model has a few limitations that should be considered when interpreting its predictions.
+First, it is trained on cross-sectional data rather than longitudinal data.
+As a result, the model is designed to provide accurate valuations for properties in the near future (typically less than a year), assuming minimal price fluctuations.
+Significant market changes over time would reduce the model's accuracy, requiring frequent updates with new cross-sectional data to maintain reliability.
+Additionally, the model is specifically tailored to Seattle’s real estate market.
+While it might perform adequately in nearby cities with similar market characteristics, it is unlikely to generalize well to cities in other states, such as Los Angeles or New York, due to differing property attributes and pricing dynamics in those regions.
+
+## Justification
+
+The four features included in the model are `beds`, `sqft`, `hoa_month`, and `sqft:beds`, as they were the most significant in influencing the model's property valuations.
+The beds feature was selected because the number of bedrooms is a key determinant of a property's utility and appeal to buyers, directly impacting its market value.
+The sqft feature, representing the total square footage, is a fundamental metric for assessing a property's size and, consequently, its worth.
+The hoa_month feature accounts for monthly homeowner association fees, which can significantly affect the affordability and desirability of properties, particularly in condominiums or communities with shared amenities.
+Finally, the interaction term sqft:beds captures the relationship between the size of the property and the number of bedrooms, highlighting how the distribution of space impacts valuation.
+
+```{r}
+full_model <- readRDS("../models/full_model.rds")
+reduced_model <- readRDS("../models/reduced_model.rds")
+final_model <- readRDS("../models/final_model.rds")
+```
+
+```{r #tbl-summ}
+modelsummary(list(
+  # "Full" = full_model,
+  # "Reduced" = reduced_model,
+  "Final Model" = final_model), output = "kableExtra",
+  gof_omit = "BIC|AIC") |>
+  column_spec(2, width = "2.5in")
+```
+
+## Interpretation
+
+@tbl-summ shows the values of the model's coefficients mentioned in the setup.
+The intercept of -341,047.42 suggests that, without any additional predictors, property prices would begin at this negative baseline, though this value has limited standalone interpretation.
+The coefficient for beds is -159,574.45, meaning that, holding other factors constant, each additional bedroom is associated with a significant reduction in price.
+However, this effect is modified by the interaction term beds $\times$ sqft, which has a negative coefficient of -32.64, indicating that larger properties with more bedrooms have diminishing price returns.
+The sqft variable has a strong positive impact (1152.06 per additional square foot), highlighting size as a key driver of value.
+The hoa_month coefficient (267.83) shows a modest positive association with price, possibly reflecting higher costs in premium communities.
+With an $R^2$ of 0.654, the model explains 65.4\% of the variance in property prices, and an RMSE of 719,452.71, the summary suggests a decent fit.
+
+## Validation
+
+The linear model was created using base R [@R] to fit the data and generate predictions, while the MLMetrics package [@mlmetrics] was used to evaluate performance.
+A train-test split of 70:30 was created with the rsample package [@rsample], with the model trained on the larger portion of the data and validated using out-of-sample testing on the remaining data.
+Key evaluation metrics included the Root Mean Square Error (RMSE) and the $R^2$ score.
+The model achieved an RMSE of 1,850,720, indicating the average prediction error in dollar terms, and an $R^2$ score of approximately 0.281, reflecting the proportion of variance in property prices explained by the model.
+Detailed diagnostics and further information are provided in the appendix.
+
+## Alternate Models
+
+Several alternative models were considered before selecting the final one.
+The first, a “full model,” included all variables in the cleaned dataset, achieving an RMSE of 1,864,966 and an $R^2$ score of 0.269.
+While this model used all available information, many variables were statistically insignificant.
+A reduced model was then tested by retaining only significant variables, resulting in an RMSE of 1,865,188 and an $R^2$ score of 0.269.
+Though concise, it performed similarly to the full model.
+The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 1,850,720 and an $R^2$ score of 0.281.
+
+# Results {#sec-results}
 
@@
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis volutpat ut ante vestibulum sagittis.
-Etiam sollicitudin ligula neque, et sollicitudin erat iaculis eu.
-Pellentesque aliquam semper sapien, a faucibus nulla consequat vitae.
-Praesent in vulputate neque.
-Fusce fermentum tempor turpis, nec iaculis diam porta nec.
-Maecenas nec porta felis, at tincidunt ex.
-Mauris convallis tincidunt posuere.
-Quisque rutrum imperdiet turpis, quis bibendum est ultrices ut.
-Phasellus in tortor eu ligula pellentesque pellentesque.
+Suspendisse accumsan elementum diam sed ullamcorper.
+Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
+Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
+Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
+Vivamus eget rutrum erat.
+Ut dictum quis ligula non euismod.
+Fusce sit amet egestas nibh.
+In rutrum tortor sed nulla volutpat facilisis.
 Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Curabitur eleifend ac purus quis dapibus.
-Ut tincidunt posuere tortor, vitae venenatis nunc placerat eu.
-Praesent sit amet mi urna.
-Sed ultrices est et ligula posuere, eget molestie lorem porta.
-Quisque tempor, dolor porta semper aliquam, urna magna rhoncus sapien, at mattis metus nibh eleifend mi.
-Sed id bibendum urna, ut tempor ligula.
-Nulla facilisi.
-Nulla facilisi.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Nunc non sapien lacus.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Fusce nec magna nec leo convallis dignissim ut porta justo.
+Curabitur nec faucibus orci.
+Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
+Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
+Morbi eget odio nisi.
+Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
 
@@
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis volutpat ut ante vestibulum sagittis.
-Etiam sollicitudin ligula neque, et sollicitudin erat iaculis eu.
-Pellentesque aliquam semper sapien, a faucibus nulla consequat vitae.
-Praesent in vulputate neque.
-Fusce fermentum tempor turpis, nec iaculis diam porta nec.
-Maecenas nec porta felis, at tincidunt ex.
-Mauris convallis tincidunt posuere.
-Quisque rutrum imperdiet turpis, quis bibendum est ultrices ut.
-Phasellus in tortor eu ligula pellentesque pellentesque.
+Suspendisse accumsan elementum diam sed ullamcorper.
+Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
+Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
+Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
+Vivamus eget rutrum erat.
+Ut dictum quis ligula non euismod.
+Fusce sit amet egestas nibh.
+In rutrum tortor sed nulla volutpat facilisis.
 Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Curabitur eleifend ac purus quis dapibus.
-Ut tincidunt posuere tortor, vitae venenatis nunc placerat eu.
-Praesent sit amet mi urna.
-Sed ultrices est et ligula posuere, eget molestie lorem porta.
-Quisque tempor, dolor porta semper aliquam, urna magna rhoncus sapien, at mattis metus nibh eleifend mi.
-Sed id bibendum urna, ut tempor ligula.
-Nulla facilisi.
-Nulla facilisi.
-Proin mollis quam ut rutrum vulputate.
-Praesent commodo.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Nunc non sapien lacus.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Fusce nec magna nec leo convallis dignissim ut porta justo.
+Curabitur nec faucibus orci.
+Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
+Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
+Morbi eget odio nisi.
+Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
 
@@
 
-# Discussion
+# Discussion {#sec-discuss}
 
@@
 
+# Additional Model Details
+
+```{r, fig.width=6, fig.height=3}
+par(mfrow = c(1, 2))
+
+plot(fitted(final_model), resid(final_model), main = "Residual-Fitted Plot",
+     xlab="Fitted", ylab = "Residual")
+qqnorm(resid(final_model));qqline(resid(final_model))
+```
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis vitae pretium velit.
+Etiam fermentum enim ac venenatis fringilla.
+Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
+Curabitur commodo pretium ornare.
+Pellentesque ultrices sapien ut lectus viverra placerat.
+Mauris nec interdum tortor.
+Cras posuere turpis sed mi blandit dictum.
+Etiam accumsan ipsum quis sem ornare cursus.
+Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
+Nullam ornare placerat risus, in luctus urna.
+Ut molestie dignissim turpis, eget tincidunt arcu.
+Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
+Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
+
+```{r}
+modelsummary(list(
+  "Full" = full_model,
+  "Reduced" = reduced_model), 
+  output = "kableExtra") |>
+  column_spec(2, width = "2in")
+```
+
 \newpage

### 2024-11-27 scripts/modeling.R
Updating models

@@ -23,12 +23,13 @@ cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
 # Preparing data for the model
 model_data <- cleaned_data |>
   select(property_type, beds, baths, half_bath, sqft, days_on_market,
-         hoa_month, property_age, price)
+         hoa_month, property_age, price) |>
+  drop_na()
 
 # Setting seed for reproducibility
 set.seed(12)
 
-split <- initial_split(model_data, prop = 0.7)
+split <- initial_split(model_data, prop = 0.7, strata = property_type)
 
 training_data <- training(split)
 testing_data <- testing(split)
@@ -41,7 +42,7 @@ summary(model1)
 
 # Fitting model 2
 model2 <- lm(data = training_data, 
-             formula = price ~ property_type + beds + sqft + property_age + hoa_month)
+             formula = price ~ beds + sqft + hoa_month)
 
 summary(model2)
 
@@ -50,29 +51,17 @@ anova(model1, model2)
 plot(fitted(model2), resid(model2))
 qqnorm(resid(model2));qqline(resid(model2))
 
-# Fitting model 3
-model3 <- lm(data = training_data, formula = price ~ beds + sqft + hoa_month + sqft:beds)
-
-summary(model3)
-
-plot(fitted(model3), resid(model3))
-qqnorm(resid(model3));qqline(resid(model3))
-
 # Evaluating models
 testing_data  <- testing_data |>
   mutate(pred1 = predict(model1, testing_data),
-         pred2 = predict(model2, testing_data),
-         pred3 = predict(model3, testing_data))
+         pred2 = predict(model2, testing_data))
 
 RMSE(testing_data$pred1, testing_data$price)
 RMSE(testing_data$pred2, testing_data$price)
-RMSE(testing_data$pred3, testing_data$price)
 
 R2_Score(testing_data$pred1, testing_data$price)
 R2_Score(testing_data$pred2, testing_data$price)
-R2_Score(testing_data$pred3, testing_data$price)
 
 # Exporting models
 saveRDS(model1, "models/full_model.rds")
 saveRDS(model2, "models/reduced_model.rds")
-saveRDS(model3, "models/final_model.rds")

### 2024-11-28 paper/paper.qmd
Fixing an error in the models

@@
 The hoa_month coefficient (267.83) shows a modest positive association with price, possibly reflecting higher costs in premium communities.
-With an $R^2$ of 0.654, the model explains 65.4\% of the variance in property prices, and an RMSE of 719,452.71, the summary suggests a decent fit.
+With an adjusted $R^2$ of 0.6543, the model explains 65.3\% of the variance in property prices, and an RMSE of 719,452.71, the summary suggests a decent fit.
 

### 2024-11-28 scripts/modeling.R

@@ -29,7 +29,7 @@ model_data <- cleaned_data |>
 # Setting seed for reproducibility
 set.seed(12)
 
-split <- initial_split(model_data, prop = 0.7, strata = property_type)
+split <- initial_split(model_data, prop = 0.7)
 
 training_data <- training(split)
 testing_data <- testing(split)
@@ -51,17 +51,30 @@ anova(model1, model2)
 plot(fitted(model2), resid(model2))
 qqnorm(resid(model2));qqline(resid(model2))
 
+# Fitting model 3
+model3 <- lm(data = training_data, 
+             formula = price ~ beds + sqft + hoa_month + sqft:beds)
+
+summary(model3)
+
+plot(fitted(model3), resid(model3))
+qqnorm(resid(model3));qqline(resid(model3))
+
 # Evaluating models
 testing_data  <- testing_data |>
   mutate(pred1 = predict(model1, testing_data),
-         pred2 = predict(model2, testing_data))
+         pred2 = predict(model2, testing_data),
+         pred3 = predict(model3, testing_data))
 
 RMSE(testing_data$pred1, testing_data$price)
 RMSE(testing_data$pred2, testing_data$price)
+RMSE(testing_data$pred3, testing_data$price)
 
 R2_Score(testing_data$pred1, testing_data$price)
 R2_Score(testing_data$pred2, testing_data$price)
+R2_Score(testing_data$pred3, testing_data$price)
 
 # Exporting models
 saveRDS(model1, "models/full_model.rds")
 saveRDS(model2, "models/reduced_model.rds")
+saveRDS(model3, "models/final_model.rds")

### 2024-11-28 scripts/cleaning.R
Updating scripts and README.md

@@ -5,9 +5,7 @@
 # Contact: sameeck.bhatia@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-#   - `tidyverse` package must be installed.
-#   - `janitor` package must be installed.
-#   - `arrow` package must be installed.
+#   - `package_installation.R` must be run.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-28 scripts/cleaning_tests.R

@@ -5,8 +5,7 @@
 # Contact: sameeck.bhatia@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-#   - `tidyverse` package must be installed.
-#   - `testthat` package must be installed.
+#   - `package_installation.R` must be run.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-28 scripts/eda.R

@@ -5,8 +5,7 @@
 # Contact: sameeck.bhatia@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-#   - `tidyverse` package must be installed.
-#   - `arrow` package must be installed.
+#   - `package_installation.R` must be run.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-28 scripts/modeling.R

@@ -5,10 +5,7 @@
 # Contact: sameeck.bhatia@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-#   - `tidyverse` package must be installed.
-#   - `arrow` package must be installed.
-#   - `rsample` package must be installed.
-#   - `MLmetrics` package must be installed.
+#   - `package_installation.R` must be run.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-28 scripts/package_installation.R

@@ -1,0 +1,17 @@
+#### Preamble ####
+# Purpose: Installs all libraries relevant to the paper.
+# Author: Sameeck Bhatia
+# Date: 27 November 2024
+# Contact: sameeck.bhatia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+# Installing packages
+install.packages("tidyverse")
+install.packages("janitor")
+install.packages("testthat")
+install.packages("rsample")
+install.packages("MLmetrics")
+install.packages("modelsummary")
+install.packages("stargazer")

### 2024-11-28 scripts/simulation.R

@@ -5,7 +5,7 @@
 # Contact: sameeck.bhatia@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-#   - `tidyverse` package must be installed.
+#   - `package_installation.R` must be run.
 # Any other information needed? None
 
 # Loading required libraries

### 2024-11-28 paper/paper.qmd
Adding and interpreting a DAG

@@
 library(modelsummary)
+library(dagitty)
+library(ggdag)
 ```
@@
 By applying machine learning and statistical techniques, the model aims to ensure accuracy and reflect the intrinsic value of properties rather than merely their market prices.
-
 Existing valuation tools, such as Zillow’s “Zestimate” and Redfin’s proprietary estimates, have limitations.
@@
 
-**TO-DO: ** Preliminary findings, significance, estimand
+**TO-DO: ** Estimand.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris rhoncus scelerisque felis, vitae feugiat est mattis non.
+Morbi blandit felis non porttitor fermentum.
+Integer arcu metus, fringilla in mattis elementum, ornare et arcu.
+Integer ut tellus id tortor pulvinar posuere.
+In molestie quam eu mi efficitur ornare.
+Fusce et egestas felis, ac rhoncus nisl.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 
+**TO-DO: ** Preliminary findings, significance.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris rhoncus scelerisque felis, vitae feugiat est mattis non.
+Morbi blandit felis non porttitor fermentum.
+Integer arcu metus, fringilla in mattis elementum, ornare et arcu.
+Integer ut tellus id tortor pulvinar posuere.
+In molestie quam eu mi efficitur ornare.
+Fusce et egestas felis, ac rhoncus nisl.
 The paper is structured as follows: @sec-data introduces the dataset and variables.
@@
 Further details about these variables can be found in Appendix A.
-The data was analyzed using `R` [@R] and the `tidyverse` [@tidyverse] package, while visualizations have been created using `tidyverse`.
+The data was analyzed using R [@R] and the tidyverse [@tidyverse] package, while visualizations have been created using tidyverse.
 
@@
 
-```{r}
-full_model <- readRDS("../models/full_model.rds")
-reduced_model <- readRDS("../models/reduced_model.rds")
-final_model <- readRDS("../models/final_model.rds")
-```
-
-```{r #tbl-summ}
-modelsummary(list(
-  # "Full" = full_model,
-  # "Reduced" = reduced_model,
-  "Final Model" = final_model), output = "kableExtra",
-  gof_omit = "BIC|AIC") |>
-  column_spec(2, width = "2.5in")
-```
-
 ## Interpretation
@@
 
+@fig-dag represents the causal relationships between the variables analyzed in the regression model.
+It illustrates that property price is influenced by the number of bedrooms, bathrooms, square footage, and HOA fees.
+Square footage itself depends on the number of bedrooms and bathrooms, while HOA fees are linked to square footage.
+The directed arrows capture the assumed causal pathways, emphasizing that the effects of bedrooms and bathrooms on price are partly mediated through square footage.
+
+```{r}
+full_model <- readRDS("../models/full_model.rds")
+reduced_model <- readRDS("../models/reduced_model.rds")
+final_model <- readRDS("../models/final_model.rds")
+```
+
+```{r #tbl-summ}
+modelsummary(list(
+  "Final Model" = final_model), output = "kableExtra",
+  gof_omit = "BIC|AIC") |>
+  column_spec(2, width = "2.5in")
+```
+
+```{r fig-dag, fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
+dagify(
+  price ~ beds,
+  price ~ sqft,
+  price ~ baths,
+  price ~ hoa,
+  sqft ~ beds,
+  sqft ~ baths,
+  baths ~ beds,
+  hoa ~ sqft,
+  coords = list(
+    x = c(beds = 0.5, sqft = 1, baths = 0, price = 2, hoa = 1.5),
+    y = c(beds = 0, sqft = 1, baths = 1, price = 0, hoa = 1.5)
+  )
+) |>
+  ggdag() +
+  theme_void()
+```
+
 ## Validation
@@
 The model achieved an RMSE of 1,850,720, indicating the average prediction error in dollar terms, and an $R^2$ score of approximately 0.281, reflecting the proportion of variance in property prices explained by the model.
-Detailed diagnostics and further information are provided in the appendix.
+Detailed diagnostics and further information are provided in Appendix B.
 
@@
 
-```{r}
+```{r #tbl-dict}
 data_dictionary <- data.frame(
@@
 
-```{r, fig.width=6, fig.height=3}
+```{r fig-diagnostics, fig.width=6, fig.height=3.5, fig.cap='Visible Violations of Constant Variance and Normality Assumptions'}
 par(mfrow = c(1, 2))
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-diagnostics ipsum dolor sit amet, consectetur adipiscing elit.
 Duis vitae pretium velit.
@@
 
-```{r}
+```{r #tbl-models}
 modelsummary(list(

### 2024-11-28 scripts/package_installation.R
Changing references and packages

@@ -8,10 +8,12 @@
 # Any other information needed? None
 
 # Installing packages
-install.packages("tidyverse")
+install.packages("daggity")
+install.packages("ggdag")
 install.packages("janitor")
-install.packages("testthat")
-install.packages("rsample")
 install.packages("MLmetrics")
 install.packages("modelsummary")
+install.packages("rsample")
 install.packages("stargazer")
+install.packages("testthat")
+install.packages("tidyverse")

### 2024-11-28 paper/paper.qmd
Adding charts

@@
 library(tidyverse)
-library(palmerpenguins)
+library(rsample)
 library(kableExtra)
@@
   ggplot(aes(x = price)) +
-  geom_histogram(binwidth = 1000000, fill = "beige", color = "black") +
+  geom_histogram(binwidth = 1000000, fill = "#519C7C", color = "black") +
   labs(x = "Price ($)", y = "Count")
@@
   ggplot(aes(x = log(price))) +
-  geom_histogram(binwidth = 0.25, fill = "beige", color = "black") +
+  geom_histogram(binwidth = 0.25, fill = "#519C7C", color = "black") +
   labs(x = " ln(Price)", y = "Count")
@@
   ggplot(aes(x = property_type)) +
-  geom_bar(fill = "beige", color = "black") +
+  geom_bar(fill = "#519C7C", color = "black") +
   labs(x = "Property Type", y = "Count")
@@
   ggplot(aes(x = beds)) +
-  geom_bar(fill = "beige", color = "black") +
+  geom_bar(fill = "#519C7C", color = "black") +
   labs(x = "Beds", y = "Count")
@@
   ggplot(aes(x = sqft)) + 
-  geom_histogram(binwidth = 500, fill = "beige", color = "black") +
+  geom_histogram(binwidth = 500, fill = "#519C7C", color = "black") +
   labs(x = "Size (sqft)", y = NULL)
@@
   ggplot(aes(x = property_age)) + 
-  geom_histogram(binwidth = 10, fill = "beige", color = "black") +
+  geom_histogram(binwidth = 10, fill = "#519C7C", color = "black") +
   labs(x = "Property Age (years)", y = NULL)
@@
 
-```{r fig-dag, fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
+```{r fig-dag, fig.align='center', fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
 dagify(
@@
 
+```{r}
+set.seed(12)
+
+split <- initial_split(data, prop = 0.7)
+
+testing_data <- testing(split)
+```
+
 ## Result 1
 
-```{r, fig.align='center'}
-palmerpenguins::penguins |>
-  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
-  geom_point()
+```{r fig-scatter, fig.cap='Large Uncertainty in the Values of Luxury Homes and Mansions'}
+testing_data |>
+  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
+  drop_na() |>
+  ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
+  geom_point(color = "#519C7C", size = 1) +
+  geom_function(fun = function(x){x}, linetype = "dashed") +
+  scale_x_continuous(breaks = seq(0, 20, 2)) +
+  scale_y_continuous(breaks = seq(0, 20, 2)) +
+  labs(x = "Predicted Price ($ millions)", y = "Actual Price ($ millions)")
 ```
@@
 
+\newpage
+
 ## Result 2
 
-```{r, fig.align='center'}
-palmerpenguins::penguins |>
+```{r fig-col, fig.cap='Highest Average Predicted Listing Price in 85 Acre Gated Community'}
+testing_data |>
+  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
   drop_na() |>
-  ggplot(aes(x = sex, y = body_mass_g, fill = sex)) +
-  geom_boxplot(width = 0.5) +
-  coord_flip()
+  select(neighbourhood, predicted_price) |>
+  group_by(neighbourhood) |>
+  summarise(mean = mean(predicted_price / 10^6)) |>
+  arrange(desc(mean)) |>
+  head(n = 5) |>
+  ggplot(aes(x = reorder(neighbourhood, mean), y = mean)) +
+  geom_col(position = "dodge", fill = "#519C7C", color = "black") +
+  scale_y_continuous(breaks = seq(0, 5, 0.5)) +
+  coord_flip() +
+  labs(x = "Neighbourhood", y = "Mean Predicted Price ($ millions)")
 ```
@@
 
+\newpage
+
 ## Result 3
 
-```{r, fig.align='center'}
-palmerpenguins::penguins |>
+```{r fig-box, fig.cap='Single Family Homes Valued Significantly Higher Than Other Property Types'}
+testing_data |>
+  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
   drop_na() |>
-  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
-  geom_violin(alpha = 0.5) +
-  geom_boxplot(width = 0.25) +
-  coord_flip()
+  ggplot(aes(x = as.character(property_type), y = predicted_price / 10^6)) +
+  geom_boxplot(fill = "#519C7C", width = 0.25) +
+  coord_flip() +
+  scale_y_continuous(breaks = seq(0, 8, 1)) +
+  labs(x = "Property Type", y = "Price ($ millions)")
 ```

### 2024-11-28 paper/paper.qmd
Adding more charts

@@
 
-```{r}
-#| output: asis
-
+```{r, output='asis'}
 data <- read_parquet("../data/cleaned_data/cleaned_data.parquet") |>
@@
   "Final Model" = final_model), output = "kableExtra",
+  fmt = "%.2f",
   gof_omit = "BIC|AIC") |>
@@
 testing_data <- testing(split)
+
+testing_data <- testing_data |>
+  mutate(predicted_price = predict(final_model, newdata = testing_data),
+         predicted_price_sqft = predicted_price / sqft) |>
+  drop_na()
 ```
@@
 testing_data |>
-  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
-  drop_na() |>
   ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
@@
 
-```{r fig-col, fig.cap='Highest Average Predicted Listing Price in 85 Acre Gated Community'}
+```{r fig-col, fig.cap="Highest Average Predicted Listing Price in Seattle's 85 Acre Gated Community"}
 testing_data |>
-  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
-  drop_na() |>
   select(neighbourhood, predicted_price) |>
@@
 
-```{r fig-box, fig.cap='Single Family Homes Valued Significantly Higher Than Other Property Types'}
+```{r fig-box, fig.cap='40 Year Old Homes Among the Lowest in Actual and Predicted Values'}
 testing_data |>
-  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
-  drop_na() |>
-  ggplot(aes(x = as.character(property_type), y = predicted_price / 10^6)) +
-  geom_boxplot(fill = "#519C7C", width = 0.25) +
-  coord_flip() +
-  scale_y_continuous(breaks = seq(0, 8, 1)) +
-  labs(x = "Property Type", y = "Price ($ millions)")
+  group_by(property_age) |>
+  summarise(mean_price = mean(price) / 10^6,
+            mean_predicted = mean(predicted_price) / 10^6) |>
+  pivot_longer(names_to = "price_type", cols = c(mean_price, mean_predicted),
+               values_to = "value") |>
+  mutate(price_type = recode(price_type, 
+                             "mean_price" = "Actual", 
+                             "mean_predicted" = "Predicted")) |>
+  ggplot(aes(x = property_age, y = value, color = price_type)) +
+  geom_point(size = 0.75) +
+  scale_color_manual(values = c("black", "#519C7C")) +
+  geom_smooth(method = "lm", formula = y ~ I(x^2) + x, se = FALSE) +
+  facet_wrap(vars(price_type)) +
+  theme(legend.position = "top") +
+  labs(x = "Property Age", y = "Average Price ($ millions)", color = "Price Type")
 ```

### 2024-11-28 scripts/modeling.R
Updating models

@@ -21,6 +21,7 @@ cleaned_data <- read_parquet("data/cleaned_data/cleaned_data.parquet")
 model_data <- cleaned_data |>
   select(property_type, beds, baths, half_bath, sqft, days_on_market,
          hoa_month, property_age, price) |>
+  mutate(price = ifelse(price < quantile(price, 0.95), price, quantile(price, 0.95))) |>
   drop_na()
 
 # Setting seed for reproducibility
@@ -50,7 +51,7 @@ qqnorm(resid(model2));qqline(resid(model2))
 
 # Fitting model 3
 model3 <- lm(data = training_data, 
-             formula = price ~ beds + sqft + hoa_month + sqft:beds)
+             formula = price ~ beds + baths + sqft + hoa_month + sqft:beds)
 
 summary(model3)
 

### 2024-11-28 paper/paper.qmd
Updating LLM usage and paper

@@
 library(ggdag)
+library(sf)
 ```
@@
 
-\newpage
-
 # Model {#sec-model}
@@
 $$
-\text{price} = \beta_0 + \beta_1 \cdot \text{beds} + \beta_2 \cdot \text{sqft} + \beta_3 \cdot \text{hoa\_month} + \beta_4 \cdot (\text{sqft} \cdot \text{beds})
+\text{price} = \beta_0 + \beta_1 \cdot \text{beds} + \beta_2 \cdot \text{baths} + \beta_3 \cdot \text{sqft} + \beta_4 \cdot \text{hoa\_month} + \beta_5 \cdot (\text{sqft} \cdot \text{beds})
 $$
@@
 
-The effect of adding one more bedroom ($\text{beds}$) on `price`, holding other factors constant (except for the interaction term).
+The effect of adding one more bedroom on price, holding other factors constant (except for the interaction term).
 
-**$\beta_2$: Coefficient for `sqft`**
+**$\beta_2$: Coefficient for `baths`**
 
-The effect of increasing square footage ($\text{sqft}$) by one unit on `price`, holding other factors constant (except for the interaction term).
+The effect of adding one more bathroom on price, holding other factors constant.
 
-**$\beta_3$: Coefficient for `hoa_month`**
+**$\beta_3$: Coefficient for `sqft`**
 
-The effect of a one-unit increase in monthly homeowner's association fees (`hoa_month`) on `price`, holding other factors constant.
+The effect of increasing square footage by one unit on price, holding other factors constant (except for the interaction term).
 
-**$\beta_4$: Interaction between `sqft` and `beds`**
+**$\beta_4$: Coefficient for `hoa_month`**
 
-Represents how the relationship between $\text{sqft}$ and `price` changes depending on the number of bedrooms ($\text{beds}$).
+The effect of a one-unit increase in monthly homeowner's association fees on price, holding other factors constant.
+
+**$\beta_5$: Interaction between `sqft` and `beds`**
+
+Represents how the relationship between $\text{sqft}$ and `price` changes depending on the number of bedrooms.
 
@@
 
-## Interpretation
-
-@tbl-summ shows the values of the model's coefficients mentioned in the setup.
-The intercept of -341,047.42 suggests that, without any additional predictors, property prices would begin at this negative baseline, though this value has limited standalone interpretation.
-The coefficient for beds is -159,574.45, meaning that, holding other factors constant, each additional bedroom is associated with a significant reduction in price.
-However, this effect is modified by the interaction term beds $\times$ sqft, which has a negative coefficient of -32.64, indicating that larger properties with more bedrooms have diminishing price returns.
-The sqft variable has a strong positive impact (1152.06 per additional square foot), highlighting size as a key driver of value.
-The hoa_month coefficient (267.83) shows a modest positive association with price, possibly reflecting higher costs in premium communities.
-With an adjusted $R^2$ of 0.6543, the model explains 65.3\% of the variance in property prices, and an RMSE of 719,452.71, the summary suggests a decent fit.
-
 @fig-dag represents the causal relationships between the variables analyzed in the regression model.
@@
 The directed arrows capture the assumed causal pathways, emphasizing that the effects of bedrooms and bathrooms on price are partly mediated through square footage.
+
+```{r fig-dag, fig.width=4, fig.height=2.8, fig.cap='Multiple Causal Relationships Among Model Variables'}
+dagify(
+  price ~ beds,
+  price ~ sqft,
+  price ~ baths,
+  price ~ hoa,
+  sqft ~ beds,
+  sqft ~ baths,
+  baths ~ beds,
+  hoa ~ sqft,
+  coords = list(
+    x = c(beds = 0.5, sqft = 0.9, baths = 0, price = 2, hoa = 1.6),
+    y = c(beds = 0, sqft = 0.9, baths = 1, price = 0, hoa = 1.2)
+  )
+) |>
+  ggdag() +
+  theme_void()
+```
+
+## Interpretation
+
+@tbl-summ shows the values of the model's coefficients mentioned in the setup.
+The intercept of −66,054.23 suggests that, without any additional predictors, property prices would begin at this value, though it has limited standalone interpretation.
+The coefficient for `beds` is 55,091.77, meaning that, holding other factors constant, each additional bedroom is associated with an addition in price.
+However, this effect is modified by the interaction term `beds` $\times$ `sqft`, which has a negative coefficient of −38.66, indicating that larger properties have diminishing prices.
+The coefficient for `baths` is 41,740.69, meaning that for each additional bathroom, the predicted value of the price increases by around \$41,740.69, holding all other variables constant.
+This suggests that bathrooms have a strong positive impact on the value of the property.
+The `sqft` variable has a strong positive impact (592.23 per additional square foot), highlighting size as a key driver of value.
+The `hoa_month` coefficient (66.39) shows a modest positive association with price, possibly reflecting higher costs in premium communities.
+With an adjusted $R^2$ of 0.716 and an RMSE of 331,563.51, the model has a decent fit.
 
@@
   "Final Model" = final_model), output = "kableExtra",
-  fmt = "%.2f",
-  gof_omit = "BIC|AIC") |>
+  fmt = "%.2f", gof_map = c("nobs", "r.squared", "adj.r.squared", "rmse")) |>
   column_spec(2, width = "2.5in")
-```
-
-```{r fig-dag, fig.align='center', fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
-dagify(
-  price ~ beds,
-  price ~ sqft,
-  price ~ baths,
-  price ~ hoa,
-  sqft ~ beds,
-  sqft ~ baths,
-  baths ~ beds,
-  hoa ~ sqft,
-  coords = list(
-    x = c(beds = 0.5, sqft = 1, baths = 0, price = 2, hoa = 1.5),
-    y = c(beds = 0, sqft = 1, baths = 1, price = 0, hoa = 1.5)
-  )
-) |>
-  ggdag() +
-  theme_void()
 ```
@@
 
-The linear model was created using base R [@R] to fit the data and generate predictions, while the MLMetrics package [@mlmetrics] was used to evaluate performance.
-A train-test split of 70:30 was created with the rsample package [@rsample], with the model trained on the larger portion of the data and validated using out-of-sample testing on the remaining data.
-Key evaluation metrics included the Root Mean Square Error (RMSE) and the $R^2$ score.
-The model achieved an RMSE of 1,850,720, indicating the average prediction error in dollar terms, and an $R^2$ score of approximately 0.281, reflecting the proportion of variance in property prices explained by the model.
-Detailed diagnostics and further information are provided in Appendix B.
+The linear model was created using R [@R] to fit the data and generate predictions, while the MLMetrics package [@mlmetrics] was used to evaluate performance.
+A train-test split was created with the rsample package [@rsample], with the model trained on the 70\% of the data and validated using out-of-sample testing on the remaining 30\%.
+Key evaluation metrics included the Root Mean Square Error (RMSE) and $R^2$ The model achieved an RMSE of 331,564, indicating the average prediction error in dollar terms, and $R^2$ of approximately 0.716, reflecting the proportion of variance in property prices explained by the model.
+Further diagnostics and information are provided in Appendix B.
 
@@
 Though concise, it performed similarly to the full model.
-The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 1,850,720 and an $R^2$ score of 0.281.
+The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 1,850,720 and an $R^2$ of 0.281.
 
@@
 
-```{r fig-scatter, fig.cap='Large Uncertainty in the Values of Luxury Homes and Mansions'}
+```{r fig-scatter, fig.cap='Large Uncertainty in the Predicted Values of Luxury Homes'}
 testing_data |>
   ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
-  geom_point(color = "#519C7C", size = 1) +
-  geom_function(fun = function(x){x}, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(0, 20, 2)) +
+  geom_point(color = "#519C7C", size = 1, shape = 15) +
+  geom_abline(linetype = "dashed") +
+  scale_x_continuous(breaks = seq(0, 20, 1)) +
   scale_y_continuous(breaks = seq(0, 20, 2)) +
-  labs(x = "Predicted Price ($ millions)", y = "Actual Price ($ millions)")
+  labs(
+    x = "Predicted Price ($ millions)",
+    y = "Actual Price ($ millions)"
+  )
 ```
@@
 
-```{r fig-col, fig.cap="Highest Average Predicted Listing Price in Seattle's 85 Acre Gated Community"}
+```{r fig-box, fig.cap="Single Family Homes Valued the Highest Among All Property Types"}
 testing_data |>
-  select(neighbourhood, predicted_price) |>
-  group_by(neighbourhood) |>
-  summarise(mean = mean(predicted_price / 10^6)) |>
-  arrange(desc(mean)) |>
-  head(n = 5) |>
-  ggplot(aes(x = reorder(neighbourhood, mean), y = mean)) +
-  geom_col(position = "dodge", fill = "#519C7C", color = "black") +
-  scale_y_continuous(breaks = seq(0, 5, 0.5)) +
+  pivot_longer(
+    cols = c(price, predicted_price), 
+    names_to = "price_type", 
+    values_to = "value"
+  ) |>
+  mutate(price_type = recode(price_type, "price" = "Actual", 
+                             "predicted_price" = "Predicted")) |>
+  ggplot(aes(x = property_type, y = value / 10^6, fill = price_type)) +
+  geom_boxplot(
+    position = position_dodge(0.6), 
+    color = "black", 
+    width = 0.5,
+    outlier.shape = 15, 
+    outlier.size = 0.5
+  ) +
+  scale_y_continuous(limits = c(0, 7)) +
   coord_flip() +
-  labs(x = "Neighbourhood", y = "Mean Predicted Price ($ millions)")
+  scale_fill_manual(values = c("grey70", "#519C7C")) +
+  theme(legend.position = "top") +
+  labs(
+    x = "Property Type", 
+    y = "Price ($ millions)", 
+    fill = "Price Type"
+  )
 ```
@@
 
-```{r fig-box, fig.cap='40 Year Old Homes Among the Lowest in Actual and Predicted Values'}
+```{r fig-age, fig.cap='40 Year Old Homes Among the Lowest in Actual and Predicted Values'}
 testing_data |>
   group_by(property_age) |>
-  summarise(mean_price = mean(price) / 10^6,
-            mean_predicted = mean(predicted_price) / 10^6) |>
-  pivot_longer(names_to = "price_type", cols = c(mean_price, mean_predicted),
-               values_to = "value") |>
+  summarise(
+    mean_price = mean(price) / 10^6,
+    mean_predicted = mean(predicted_price) / 10^6
+  ) |>
+  pivot_longer(
+    names_to = "price_type", 
+    cols = c(mean_price, mean_predicted),
+    values_to = "value"
+  ) |>
   mutate(price_type = recode(price_type, 
@@
   ggplot(aes(x = property_age, y = value, color = price_type)) +
-  geom_point(size = 0.75) +
+  geom_point(size = 0.75, shape = 15) +
   scale_color_manual(values = c("black", "#519C7C")) +
@@
   facet_wrap(vars(price_type)) +
-  theme(legend.position = "top") +
-  labs(x = "Property Age", y = "Average Price ($ millions)", color = "Price Type")
+  theme(
+    legend.position = "top"
+  ) +
+  labs(
+    x = "Property Age",
+    y = "Average Price ($ millions)",
+    color = "Price Type"
+  )
+```
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Suspendisse accumsan elementum diam sed ullamcorper.
+Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
+Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
+Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
+Vivamus eget rutrum erat.
+Ut dictum quis ligula non euismod.
+Fusce sit amet egestas nibh.
+In rutrum tortor sed nulla volutpat facilisis.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Nunc non sapien lacus.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Fusce nec magna nec leo convallis dignissim ut porta justo.
+Curabitur nec faucibus orci.
+Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
+Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
+Morbi eget odio nisi.
+Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
+
+\newpage
+
+## Result 4
+
+```{r fig-map, fig.cap="Highest Average Predicted Listing Price in Seattle's 85 Acre Gated Community"}
+zip_codes <- st_read("../data/spatial_data/zip-codes.geojson", quiet = TRUE) |>
+  rename(zip_code = ZCTA5CE10) |> 
+  select(zip_code, geometry)
+
+testing_data |>
+  group_by(zip_code) |>
+  summarise(
+    mean_price = mean(price) / 10^6,
+    mean_predicted = mean(predicted_price) / 10^6,
+    .groups = "drop"
+  ) |>
+  pivot_longer(
+    names_to = "price_type", 
+    cols = c(mean_price, mean_predicted),
+    values_to = "value"
+  ) |>
+  mutate(price_type = recode(price_type, 
+                             "mean_price" = "Actual", 
+                             "mean_predicted" = "Predicted")) |>
+  merge(zip_codes, by = "zip_code") |>
+  st_sf() |>  # Convert to sf object
+  ggplot() +
+  geom_sf(aes(fill = value)) +
+  coord_sf(
+    xlim = c(-122.5369, -122.1166), 
+    ylim = c(47.44308, 47.79792),
+    expand = FALSE
+  ) +
+  scale_fill_distiller(palette = "Greens", direction = 1) +
+  facet_wrap(vars(price_type)) +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank()
+  ) +
+  labs(fill = "Price \n($ millions)")
 ```

### 2024-12-01 paper/paper.qmd
Adding the estimand and updating the model

@@
 
-**TO-DO: ** Estimand.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris rhoncus scelerisque felis, vitae feugiat est mattis non.
-Morbi blandit felis non porttitor fermentum.
-Integer arcu metus, fringilla in mattis elementum, ornare et arcu.
-Integer ut tellus id tortor pulvinar posuere.
-In molestie quam eu mi efficitur ornare.
-Fusce et egestas felis, ac rhoncus nisl.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+The primary estimand of this paper is the intrinsic value of a property.
+This value is determined using a linear regression model applied to observational data collected across Seattle.
+The analysis aims to identify the true value of a property that a buyer should consider paying and a seller should consider accepting.
+Additionally, it examines the main factors influencing a property’s intrinsic value and allows for comparison to market values.
 
@@
 **Number of Bedrooms (`beds`):** This numeric variable represents the count of full bedrooms in a property, as measured by appraisers.
+It is a key factor in the analysis, as more bedrooms often correlate with greater living space and the potential to accommodate larger households.
+
+**TO-DO:** Number of Bedrooms (`beds`): This numeric variable represents the count of full bedrooms in a property, as measured by appraisers.
 It is a key factor in the analysis, as more bedrooms often correlate with greater living space and the potential to accommodate larger households.
@@
 
-```{r fig-dag, fig.width=4, fig.height=2.8, fig.cap='Multiple Causal Relationships Among Model Variables'}
+```{r fig-dag, fig.width=4, fig.height=3, fig.cap='Multiple Causal Relationships Among Model Variables'}
 dagify(
@@
   coords = list(
-    x = c(beds = 0.5, sqft = 0.9, baths = 0, price = 2, hoa = 1.6),
-    y = c(beds = 0, sqft = 0.9, baths = 1, price = 0, hoa = 1.2)
+    x = c(beds = 0.5, sqft = 1, baths = 0, price = 2, hoa = 1.5),
+    y = c(beds = 0, sqft = 1, baths = 1, price = 0, hoa = 1.5)
   )
@@
 A train-test split was created with the rsample package [@rsample], with the model trained on the 70\% of the data and validated using out-of-sample testing on the remaining 30\%.
-Key evaluation metrics included the Root Mean Square Error (RMSE) and $R^2$ The model achieved an RMSE of 331,564, indicating the average prediction error in dollar terms, and $R^2$ of approximately 0.716, reflecting the proportion of variance in property prices explained by the model.
+Key evaluation metrics included the Root Mean Square Error (RMSE) and $R^2$ The model achieved an RMSE of 331,564, indicating the average prediction error in dollar terms, and $R^2$ of approximately 0.717, reflecting the proportion of variance in property prices explained by the model.
 Further diagnostics and information are provided in Appendix B.
@@
 Several alternative models were considered before selecting the final one.
-The first, a “full model,” included all variables in the cleaned dataset, achieving an RMSE of 1,864,966 and an $R^2$ score of 0.269.
-While this model used all available information, many variables were statistically insignificant.
-A reduced model was then tested by retaining only significant variables, resulting in an RMSE of 1,865,188 and an $R^2$ score of 0.269.
+The first, a “full model,” included all variables in the cleaned dataset, achieving an RMSE of 333,693.77 and an $R^2$ score of 0.714.
+While this model used all available information, some variables were statistically insignificant.
+A reduced model was then tested by retaining only significant variables, resulting in an RMSE of 333,736.71 and an $R^2$ score of 0.713.
 Though concise, it performed similarly to the full model.
-The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 1,850,720 and an $R^2$ of 0.281.
+The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 331 563.51 and an $R^2$ of 0.717.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-scatter ipsum dolor sit amet, consectetur adipiscing elit.
 Suspendisse accumsan elementum diam sed ullamcorper.
@@
 Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
-
-\newpage
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-box ipsum dolor sit amet, consectetur adipiscing elit.
 Suspendisse accumsan elementum diam sed ullamcorper.
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-age ipsum dolor sit amet, consectetur adipiscing elit.
 Suspendisse accumsan elementum diam sed ullamcorper.
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+@fig-map ipsum dolor sit amet, consectetur adipiscing elit.
 Suspendisse accumsan elementum diam sed ullamcorper.
@@
 
-plot(fitted(final_model), resid(final_model), main = "Residual-Fitted Plot",
+plot(fitted(final_model), resid(final_model), cex = 0.5,
+     main = "Residual-Fitted Plot",
      xlab="Fitted", ylab = "Residual")
-qqnorm(resid(final_model));qqline(resid(final_model))
+qqnorm(resid(final_model), cex = 0.5);qqline(resid(final_model))
 ```
@@
 @fig-diagnostics ipsum dolor sit amet, consectetur adipiscing elit.
+Duis vitae pretium velit.
+Etiam fermentum enim ac venenatis fringilla.
+Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
+Curabitur commodo pretium ornare.
+Pellentesque ultrices sapien ut lectus viverra placerat.
+Mauris nec interdum tortor.
+Cras posuere turpis sed mi blandit dictum.
+Etiam accumsan ipsum quis sem ornare cursus.
+Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
+Nullam ornare placerat risus, in luctus urna.
+Ut molestie dignissim turpis, eget tincidunt arcu.
+Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
+Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
+
+@tbl-models ipsum dolor sit amet, consectetur adipiscing elit.
 Duis vitae pretium velit.

### 2024-12-01 scripts/modeling.R

@@ -40,7 +40,7 @@ summary(model1)
 
 # Fitting model 2
 model2 <- lm(data = training_data, 
-             formula = price ~ beds + sqft + hoa_month)
+             formula = price ~ . - beds - baths)
 
 summary(model2)
 

### 2024-12-01 scripts/package_installation.R

@@ -14,6 +14,7 @@ install.packages("janitor")
 install.packages("MLmetrics")
 install.packages("modelsummary")
 install.packages("rsample")
+install.packages("sf")
 install.packages("stargazer")
 install.packages("testthat")
 install.packages("tidyverse")

### 2024-12-01 paper/paper.qmd
Updating format and results section

@@
 ---
-title: "Valuing Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach"
-subtitle: "Subtitle"
+title: "Capturing the True Value of Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach"
+subtitle: "Location and Age Prove Key Determinants in Calculating Intrinsic Value"
 author:
@@
 thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
-abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+abstract: "**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Morbi faucibus ipsum dolor, in tincidunt justo tempor sit amet proin non turpis a orci imperdiet pretium.
@@
 
-**TO-DO: ** Preliminary findings, significance.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris rhoncus scelerisque felis, vitae feugiat est mattis non.
-Morbi blandit felis non porttitor fermentum.
-Integer arcu metus, fringilla in mattis elementum, ornare et arcu.
-Integer ut tellus id tortor pulvinar posuere.
-In molestie quam eu mi efficitur ornare.
-Fusce et egestas felis, ac rhoncus nisl.
+Preliminary findings indicate that the model effectively predicts property prices, though deviations occur at extreme price ranges.
+These discrepancies suggest that additional factors, such as high-end amenities and neighborhood characteristics, play a significant role in influencing property value.
+Understanding these dynamics is important in incorporating socio-economic and structural variables into property valuation models, ultimately improving their accuracy and real-world applicability.
 The paper is structured as follows: @sec-data introduces the dataset and variables.
@@
 
-The primary goal of this paper is to estimate the outcome variable, price, using a valuation model.
-This variable represents the market value of active listings at the time of data collection, focusing exclusively on properties located in Seattle.
-@fig-out provides insights into the distribution of property prices.
-Plot A shows that property prices are highly skewed, with a maximum value near \$40,000,000 and several listings exceeding \$5,000,000.
-To better understand the distribution, Plot B presents the logarithmic transformation of property prices.
-While the transformed distribution is less extreme, it remains skewed, indicating the presence of properties with exceptionally high valuations in the Seattle real estate market.
-
 ```{r fig-out, fig.width=7, fig.height=2.5, fig.cap='Heavily Skewed Property Prices and Skewed Still for Logarithmic Prices'}
@@
 
+The primary goal of this paper is to estimate the outcome variable, price, using a valuation model.
+This variable represents the market value of active listings at the time of data collection, focusing exclusively on properties located in Seattle.
+@fig-out provides insights into the distribution of property prices.
+Plot A shows that property prices are highly skewed, with a maximum value near \$40,000,000 and several listings exceeding \$5,000,000.
+To better understand the distribution, Plot B presents the logarithmic transformation of property prices.
+While the transformed distribution is less extreme, it remains skewed, indicating the presence of properties with exceptionally high valuations in the Seattle real estate market.
+
 ## Predictor Variables
@@
 
-**TO-DO:** Number of Bedrooms (`beds`): This numeric variable represents the count of full bedrooms in a property, as measured by appraisers.
-It is a key factor in the analysis, as more bedrooms often correlate with greater living space and the potential to accommodate larger households.
+**TO-DO:** Number of Bathrooms (`baths`): This numeric variable represents the count of full bathrooms in a property, as measured by appraisers.
+It is a key factor in the analysis, as more bathrooms often correlate with greater living space and the potential to accommodate larger households.
 
@@
 Age is a vital consideration in determining price, as newer properties are often smaller due to rising construction costs and increasingly strict zoning regulations.
-
-@fig-pred1 represents the distributions of property types and number of bedrooms.
-Plot C shows that single-family homes are the most common, followed closely by condos, with townhouses being the least prevalent.
-This likely reflects the mix of apartments and houses present in Seattle's real estate market.
-Plot D highlights that properties with 2 or 3 bedrooms are most frequent, while some properties even have north of 7 bedrooms.
-This is likely because condos typically feature 1-2 bedrooms, whereas single-family homes and townhouses commonly have 3-4 bedrooms.
 
@@
 
-@fig-pred2 displays the distributions of property size and age.
-Plot E shows that property size is positively skewed, with most properties under 2,500 square feet and a peak between 700 and 1,200 square feet.
-Properties exceeding 5,000 square feet likely represent luxury homes or mansions.
-Plot F indicates that property age is also positively skewed, with the most common properties being less than 10 years old.
-The second most common group is 20-30 years old, likely reflecting construction surges before the 2008 financial crisis and real estate market downturn.
+@fig-pred1 represents the distributions of property types and number of bedrooms.
+Plot C shows that single-family homes are the most common, followed closely by condos, with townhouses being the least prevalent.
+This likely reflects the mix of apartments and houses present in Seattle's real estate market.
+Plot D highlights that properties with 2 or 3 bedrooms are most frequent, while some properties even have north of 7 bedrooms.
+This is likely because condos typically feature 1-2 bedrooms, whereas single-family homes and townhouses commonly have 3-4 bedrooms.
 
@@
 ```
+
+@fig-pred2 displays the distributions of property size and age.
+Plot E shows that property size is positively skewed, with most properties under 2,500 square feet and a peak between 700 and 1,200 square feet.
+Properties exceeding 5,000 square feet likely represent luxury homes or mansions.
+Plot F indicates that property age is also positively skewed, with the most common properties being less than 10 years old.
+The second most common group is 20-30 years old, likely reflecting construction surges before the 2008 financial crisis and real estate market downturn.
 
@@
 
-## Result 1
+## Higher Prices Associated with Higher Prediction Errors
 
@@
 
-@fig-scatter ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse accumsan elementum diam sed ullamcorper.
-Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
-Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
-Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
-Vivamus eget rutrum erat.
-Ut dictum quis ligula non euismod.
-Fusce sit amet egestas nibh.
-In rutrum tortor sed nulla volutpat facilisis.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc non sapien lacus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Fusce nec magna nec leo convallis dignissim ut porta justo.
-Curabitur nec faucibus orci.
-Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
-Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
-Morbi eget odio nisi.
-Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
+@fig-scatter shows the model-predicted prices alongside the actual price observations from the data.
+The dashed line represents the line of equality, where the predicted price matches the actual property price.
+Some data points deviate from this line, particularly at lower values, which is expected since no model can perfectly capture all variations in the data.
+In this case, the deviations may result from missing factors that could enhance the property valuation model.
+However, when property values exceed approximately 2 million dollars, the actual prices deviate significantly from the line of equality.
+This indicates higher prediction errors, likely due to a combination of internal and external factors.
+Internal factors might include additional amenities (e.g., pools, large gardens, or extra rooms) or premium furnishings (e.g., expensive flooring or tiling), which increase lot size or property desirability, driving up prices.
+External factors may include the property’s location in high-income neighborhoods or proximity to desirable features such as beaches, shopping malls, or tourist attractions.
+These influences are analyzed further in @fig-map.
 
-## Result 2
+## Similar Distributions to Market Price from the Model
 
@@
 
-@fig-box ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse accumsan elementum diam sed ullamcorper.
-Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
-Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
-Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
-Vivamus eget rutrum erat.
-Ut dictum quis ligula non euismod.
-Fusce sit amet egestas nibh.
-In rutrum tortor sed nulla volutpat facilisis.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc non sapien lacus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Fusce nec magna nec leo convallis dignissim ut porta justo.
-Curabitur nec faucibus orci.
-Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
-Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
-Morbi eget odio nisi.
-Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
+@fig-box illustrates the distributions of actual and predicted prices based on MLS-classified property types.
+Condos have the lowest median prices, followed by townhouses, with single-family homes being the most expensive.
+This pricing trend is primarily influenced by the size of each property type, which plays a significant role in determining value.
+The predicted and actual distributions for all property types show similar shapes, although there are differences in their centers and outliers.
+The model predicts slightly higher intrinsic values for townhouses and condos compared to the actual prices, whereas there is a notable difference in the median price for single-family homes.
+This discrepancy is likely due to the real estate market being in a cooling phase, with prices remaining relatively stable over the past six months.
+Additionally, while market prices are subject to short-term fluctuations, intrinsic values tend to remain more consistent.
 
-## Result 3
+## Influence of the Time Effect on Property Valuation
 
@@
 
-@fig-age ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse accumsan elementum diam sed ullamcorper.
-Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
-Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
-Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
-Vivamus eget rutrum erat.
-Ut dictum quis ligula non euismod.
-Fusce sit amet egestas nibh.
-In rutrum tortor sed nulla volutpat facilisis.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc non sapien lacus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Fusce nec magna nec leo convallis dignissim ut porta justo.
-Curabitur nec faucibus orci.
-Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
-Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
-Morbi eget odio nisi.
-Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
+@fig-age highlights the trend between property age and the predicted and actual average prices.
+The relationship can be divided into three stages.
+First, the average price for brand-new properties starts at around 1 million dollars, followed by a gradual decrease as the properties age.
+This decline is expected, as older properties typically lose value due to higher upkeep costs and reduced desirability.
+Next, the average price bottoms out when properties reach around 40 years of age.
+This likely represents a threshold where property age significantly affects market value (for actual prices) or true value (for predicted prices).
+In the third stage, average prices increase for properties over 50 years old, continuing to rise for homes built a century ago.
+While this may seem counterintuitive given the earlier trend, it likely reflects the premium placed on vintage properties.
+Such homes often have historical significance, unique architectural features, or are built on highly desirable land, all of which enhance their perceived and intrinsic value.
 
-## Result 4
+## Noticeable Clusters for High-End Properties
 
-```{r fig-map, fig.cap="Highest Average Predicted Listing Price in Seattle's 85 Acre Gated Community"}
+```{r fig-map, fig.cap="Highest Predicted Listing Prices in Seattle's Largest Gated Communities"}
 zip_codes <- st_read("../data/spatial_data/zip-codes.geojson", quiet = TRUE) |>
@@
   summarise(
-    mean_price = mean(price) / 10^6,
-    mean_predicted = mean(predicted_price) / 10^6,
+    mean_price = median(price) / 10^6,
+    mean_predicted = median(predicted_price) / 10^6,
     .groups = "drop"
@@
   ) +
-  labs(fill = "Price \n($ millions)")
+  labs(fill = "Median Price \n($ millions)")
 ```
 
-@fig-map ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse accumsan elementum diam sed ullamcorper.
-Integer sodales posuere mauris, sit amet blandit tortor scelerisque non.
-Nulla ut sapien vel orci scelerisque varius malesuada vel lorem.
-Integer sit amet enim mattis nisi porttitor sodales nec vitae odio.
-Vivamus eget rutrum erat.
-Ut dictum quis ligula non euismod.
-Fusce sit amet egestas nibh.
-In rutrum tortor sed nulla volutpat facilisis.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc non sapien lacus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Fusce nec magna nec leo convallis dignissim ut porta justo.
-Curabitur nec faucibus orci.
-Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum.
-Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas.
-Morbi eget odio nisi.
-Curabitur posuere scelerisque libero, non pharetra ex scelerisque.
+@fig-map presents the spatial distribution of median property values across Seattle, categorized by ZIP code.
+Two distinct patterns emerge from both maps: first, two ZIP codes stand out with the highest median listing prices; second, property values decrease the farther south they are in the city.
+The model captures these patterns well, showing similar clusters.
+The two ZIP codes with the darkest shades correspond to the Broadmoor and Broadview neighborhoods, which are 85- and 1,100-acre gated communities with affluent residents and amenities such as golf courses.
+These communities share features like higher incomes, better security, and a higher standard of living, enabling them to support significantly higher property values.
+Conversely, neighborhoods in southern Seattle tend to have lower incomes, higher crime rates, and less development, and evidently lower property values.
+These factors collectively explain the disparities in property values across the city.
+These disparities reflect broader socio-economic trends in urban areas, where wealthier neighborhoods often attract more investment in infrastructure, public services, and amenities, driving property values higher.
 
@@
 
-## Discussion 1
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+## Examining Racial Bias in Real Estate Price Estimates
+
+**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
@@
 
-## Discussion 2
+## Using Alternative Algorithms for Nonlinear Price Trends
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
@@
 
-## Discussion 3
+## Including External Factors into Property Valuation Models
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
@@
 # Additional Data Details
+
+## Raw Data Dictionary
 
@@
 
+## Model Diagnostics
+
 ```{r fig-diagnostics, fig.width=6, fig.height=3.5, fig.cap='Visible Violations of Constant Variance and Normality Assumptions'}
@@
 
-@fig-diagnostics ipsum dolor sit amet, consectetur adipiscing elit.
+**TO-DO:** @fig-diagnostics ipsum dolor sit amet, consectetur adipiscing elit.
 Duis vitae pretium velit.
@@
 
-@tbl-models ipsum dolor sit amet, consectetur adipiscing elit.
+\newpage
+
+## Alternate Model Summaries
+
+**TO-DO:** @tbl-models ipsum dolor sit amet, consectetur adipiscing elit.
 Duis vitae pretium velit.
@@
 
+# Surveys, Sampling, and Observational Data
+
+**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Fusce eu hendrerit mauris, in sodales mi.
+Aenean eget urna rutrum, tristique lorem vel, maximus ipsum.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Nunc eget rutrum est.
+Donec tempor elit non nulla cursus, in consectetur est ullamcorper.
+Nam vitae porttitor felis, tempor maximus leo.
+Aenean ultricies molestie facilisis.
+Donec eget gravida eros, sit amet ornare erat.
+Cras a vehicula nibh.
+Nunc ac ligula malesuada, bibendum ante eu, luctus felis.
+Proin in tellus eu diam aliquam ultricies.
+
+Duis sit amet augue sagittis, commodo eros sollicitudin, congue libero.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Maecenas convallis erat velit, eu consequat massa posuere et.
+Nullam tempor purus sit amet quam lacinia euismod.
+Praesent sagittis risus sit amet odio porttitor, eget pellentesque turpis luctus.
+Proin feugiat convallis diam, id vulputate mauris.
+Nunc facilisis quam nec tristique pellentesque.
+Sed laoreet tortor at ipsum faucibus iaculis.
+Sed cursus convallis molestie.
+Aliquam et accumsan tortor.
+Morbi volutpat rutrum elit sed fermentum.
+In vestibulum placerat eros, et consectetur lacus fringilla a.
+Mauris elementum velit vel eros placerat, eget commodo nisi convallis.
+Sed ac nisi mi.
+Nullam malesuada auctor semper.
+
+Vestibulum at tortor ipsum.
+Quisque a elit ut odio suscipit fringilla ac ut nunc.
+Vivamus eget neque vel est iaculis sollicitudin in et turpis.
+Pellentesque quis risus ullamcorper metus porta viverra.
+In hac habitasse platea dictumst.
+In faucibus, sapien sed placerat semper, sapien lectus placerat leo, sit amet tempus risus dolor non erat.
+Morbi ac urna venenatis, pellentesque odio vel, iaculis tellus.
+Etiam tincidunt fringilla suscipit.
+Donec fermentum sollicitudin metus, quis tempor leo elementum eu.
+
 \newpage

### 2024-12-03 paper/paper.qmd
Completing abstract and discussion sections

@@
 ---
-title: "Capturing the True Value of Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach"
+title: "Capturing the True Value of Real Estate in Seattle Through Property Characteristics: A Machine Learning Approach"
 subtitle: "Location and Age Prove Key Determinants in Calculating Intrinsic Value"
@@
 thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
-abstract: "**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi faucibus ipsum dolor, in tincidunt justo tempor sit amet proin non turpis a orci imperdiet pretium.
-Aenean aliquam eros eu consectetur cursus.
-Proin vel tempor ante vestibulum felis nisi, ultricies sed porttitor in, viverra nec augue.
-Sed convallis a tortor in pretium.
-Vivamus metus ipsum, faucibus vitae turpis sed, sodales auctor dolor sed a orci id orci pulvinar tincidunt sed ut tellus.
-Morbi venenatis tempor arcu, a auctor dolor congue sit amet.
-Nam auctor tincidunt ligula rutrum aliquam.
-Sed facilisis at nulla et posuere."
+abstract: "This paper develops a transparent property valuation model for Seattle’s real estate market, using Redfin data and statistical techniques to estimate intrinsic property values.
+The model highlights key factors influencing prices, including amenities, neighborhood characteristics, and socio-economic trends, offering an accessible tool for informed decision-making.
+Results show higher prediction errors for high-value properties and reveal time-related trends, with vintage homes gaining value due to historical significance.
+Spatial analysis captures disparities in property values, reflecting income and infrastructure differences across neighborhoods."
 toc: true
@@
 Seattle’s real estate market has experienced significant fluctuations since the start of the millennium.
-Despite these booms and busts, property prices in Seattle remain among the highest in the United States.
+Despite these booms and busts, property prices in Seattle remain among the highest in the United States [@jones].
 Demand has steadily increased, and even during periods of lower demand, the market continues to attract buyers.
@@
 By applying machine learning and statistical techniques, the model aims to ensure accuracy and reflect the intrinsic value of properties rather than merely their market prices.
-Existing valuation tools, such as Zillow’s “Zestimate” and Redfin’s proprietary estimates, have limitations.
+Existing valuation tools, such as Zillow’s “Zestimate” [@zillow] and Redfin’s proprietary estimates, have limitations.
 These models are often closed-source, leaving potential users unable to access the methodology or even the results without cost.
@@
 **Property Type (`property_type`):** This categorical variable identifies the type of property for each observation, based on classifications set by the MLS.
-The three main property types included in the analysis are "Condo/Co-op," "Single Family," and "Townhouse." These classifications are critical for valuation, as different property types possess distinct features that influence their market value.
+The three main property types included in the analysis are "Condo/Co-op," "Single Family," and "Townhouse." These classifications are important for valuation, as different property types possess distinct features that influence their market value.
 
@@
 
-**TO-DO:** Number of Bathrooms (`baths`): This numeric variable represents the count of full bathrooms in a property, as measured by appraisers.
-It is a key factor in the analysis, as more bathrooms often correlate with greater living space and the potential to accommodate larger households.
+**Number of Bathrooms (`baths`):** This variable represents the total count of full bathrooms in a property, as measured by appraisers.
+Properties with more bathrooms typically appeal to larger households, and is sometimes caused by the number of bedrooms in a property, leading to correlated prices.
 
@@
 This indicates higher prediction errors, likely due to a combination of internal and external factors.
-Internal factors might include additional amenities (e.g., pools, large gardens, or extra rooms) or premium furnishings (e.g., expensive flooring or tiling), which increase lot size or property desirability, driving up prices.
+Internal factors might include additional amenities (e.g., pools, large gardens, or extra rooms) or premium furnishings (e.g., expensive flooring or tiling), which increase lot size or property desirability [@narPool], driving up prices.
 External factors may include the property’s location in high-income neighborhoods or proximity to desirable features such as beaches, shopping malls, or tourist attractions.
@@
 Condos have the lowest median prices, followed by townhouses, with single-family homes being the most expensive.
-This pricing trend is primarily influenced by the size of each property type, which plays a significant role in determining value.
+This pricing trend is primarily influenced by the size of each property type, which plays a significant role in determining value [@apartmentlist].
 The predicted and actual distributions for all property types show similar shapes, although there are differences in their centers and outliers.
 The model predicts slightly higher intrinsic values for townhouses and condos compared to the actual prices, whereas there is a notable difference in the median price for single-family homes.
-This discrepancy is likely due to the real estate market being in a cooling phase, with prices remaining relatively stable over the past six months.
+This discrepancy is likely due to the real estate market being in a cooling phase [@weisbarth], with prices remaining relatively stable over the past six months.
 Additionally, while market prices are subject to short-term fluctuations, intrinsic values tend to remain more consistent.
@@
 First, the average price for brand-new properties starts at around 1 million dollars, followed by a gradual decrease as the properties age.
-This decline is expected, as older properties typically lose value due to higher upkeep costs and reduced desirability.
+This decline is expected, as older properties typically lose value due to higher upkeep costs and reduced desirability [@coulson].
 Next, the average price bottoms out when properties reach around 40 years of age.
@@
 These factors collectively explain the disparities in property values across the city.
-These disparities reflect broader socio-economic trends in urban areas, where wealthier neighborhoods often attract more investment in infrastructure, public services, and amenities, driving property values higher.
+These disparities reflect broader socio-economic trends in urban areas, where wealthier neighborhoods often attract more investment in infrastructure, public services, and amenities, driving property values higher [@kim].
 
@@
 
-**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
-Curabitur ac leo non nunc ultrices dignissim.
-Phasellus faucibus elit id nulla pretium volutpat.
-Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo.
-Curabitur vulputate imperdiet nulla eu finibus.
-Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam.
-Etiam nunc libero, faucibus nec porta a, dignissim a nisi.
-Phasellus laoreet elementum turpis a vestibulum.
-Quisque in urna varius, lobortis eros eu, blandit sapien.
-In lobortis orci id nunc facilisis, non tempor nisl ultricies.
-Sed pellentesque tortor cursus ipsum fringilla interdum.
-Cras lacinia tempor placerat.
-Proin eu elit elit.
-Sed molestie dolor et posuere sodales.
+The integration of algorithmic estimates like Zillow’s Zestimate into real estate markets offers potential to reduce racial biases in property valuation [@yu].
+Historically, systemic practices such as redlining have resulted in undervaluation of properties in minority neighborhoods, reinforcing economic disparities.
+The Zestimate, which aggregates data from multiple sources and provides consistent valuations across neighborhoods, appears less influenced by these historical biases.
+As a public and standardized information source, it helps market participants align on property values, leading to reduced racial disparities in final sale prices compared to initial list prices.
+This effect is especially meaningful in addressing the “white premium” often observed in traditional valuation methods.
 
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-Donec at malesuada purus.
-Cras ultrices vitae metus eget aliquet.
-Aliquam sollicitudin elit placerat, porta mauris in, semper felis.
-Nulla vestibulum dolor sed tempor efficitur.
-In bibendum libero ex, in sollicitudin arcu maximus porttitor.
-Pellentesque ac sapien sit amet magna ornare auctor.
-In hac habitasse platea dictumst.
-Cras ut urna dapibus elit mattis aliquam non sed augue.
-Suspendisse lacinia ornare consectetur.
-Quisque ultricies tortor non ullamcorper porta.
-Mauris ac leo quis lacus tincidunt cursus ut id felis.
+However, while the Zestimate mitigates some biases, it does not entirely eliminate racial inequities in the housing market.
+Factors like historical undervaluation and socio-economic disparities remain deeply embedded, requiring more targeted interventions beyond algorithmic tools.
+Moreover, reliance on algorithmic estimates can sometimes oversimplify the nuanced factors that influence property values, such as cultural significance or community dynamics, which are more difficult to quantify.
+Nevertheless, the use of transparent and publicly available valuation tools represents a positive step toward a more equitable real estate market.
 
@@
 
-**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
-Curabitur ac leo non nunc ultrices dignissim.
-Phasellus faucibus elit id nulla pretium volutpat.
-Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo.
-Curabitur vulputate imperdiet nulla eu finibus.
-Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam.
-Etiam nunc libero, faucibus nec porta a, dignissim a nisi.
-Phasellus laoreet elementum turpis a vestibulum.
-Quisque in urna varius, lobortis eros eu, blandit sapien.
-In lobortis orci id nunc facilisis, non tempor nisl ultricies.
-Sed pellentesque tortor cursus ipsum fringilla interdum.
-Cras lacinia tempor placerat.
-Proin eu elit elit.
-Sed molestie dolor et posuere sodales.
+Machine learning techniques have become invaluable for modeling non-linear trends in property prices, capturing the complex relationships between property characteristics, location, and market factors.
+Traditional methods, such as linear price models, often struggle with these complexities, especially when spatial dependencies at regional levels influence property values.
+Advanced algorithms like Random Forest and Gradient Boosting have shown promise in addressing these challenges.
+By using data at finer geographical scales, such as statistical sub-areas, these models achieve greater accuracy, particularly when predicting price differences between housing types like apartments and standalone homes.
+This spatially informed approach helps stakeholders understand localized price trends and refine policy-making [@gao].
 
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-Donec at malesuada purus.
-Cras ultrices vitae metus eget aliquet.
-Aliquam sollicitudin elit placerat, porta mauris in, semper felis.
-Nulla vestibulum dolor sed tempor efficitur.
-In bibendum libero ex, in sollicitudin arcu maximus porttitor.
-Pellentesque ac sapien sit amet magna ornare auctor.
-In hac habitasse platea dictumst.
-Cras ut urna dapibus elit mattis aliquam non sed augue.
-Suspendisse lacinia ornare consectetur.
-Quisque ultricies tortor non ullamcorper porta.
-Mauris ac leo quis lacus tincidunt cursus ut id felis.
+Emerging methods like Artificial Neural Networks and Fuzzy Inference Systems further enhance the ability to capture non-linear patterns in real estate data.
+Recent developments in fuzzy regression techniques have demonstrated superior performance in predicting prices while maintaining lower computational costs compared to older models.
+These approaches excel at interpreting complex interactions between pricing attributes and property values, offering improved precision even for less structured data like textual property descriptions.
+Such innovations highlight the growing potential of machine learning to address long-standing challenges in real estate valuation, providing better tools for buyers, sellers, and policymakers to navigate market dynamics [@sarip].
 
@@
 
-**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh.
-Curabitur ac leo non nunc ultrices dignissim.
-Phasellus faucibus elit id nulla pretium volutpat.
-Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo.
-Curabitur vulputate imperdiet nulla eu finibus.
-Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam.
-Etiam nunc libero, faucibus nec porta a, dignissim a nisi.
-Phasellus laoreet elementum turpis a vestibulum.
-Quisque in urna varius, lobortis eros eu, blandit sapien.
-In lobortis orci id nunc facilisis, non tempor nisl ultricies.
-Sed pellentesque tortor cursus ipsum fringilla interdum.
-Cras lacinia tempor placerat.
-Proin eu elit elit.
-Sed molestie dolor et posuere sodales.
+Incorporating external environmental factors into property valuation models offers valuable advantages, such as a more comprehensive understanding of real estate prices influenced by location-specific qualities [@din] [@abelson].
+Geographic information systems (GIS) provide detailed data on factors like neighborhood quality, proximity to amenities, and environmental conditions, enabling models to quantify their impact on property values.
+For example, positive attributes such as good views or access to shops can enhance valuation, while disamenities like noise pollution may reduce it.
+These factors add depth to traditional models, ensuring they reflect localized preferences and market trends, which is particularly helpful for urban planning and decision-making.
 
-Fusce aliquam tempus tincidunt.
-Nulla viverra non nibh consectetur blandit.
-Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus.
-Donec at malesuada purus.
-Cras ultrices vitae metus eget aliquet.
-Aliquam sollicitudin elit placerat, porta mauris in, semper felis.
-Nulla vestibulum dolor sed tempor efficitur.
-In bibendum libero ex, in sollicitudin arcu maximus porttitor.
-Pellentesque ac sapien sit amet magna ornare auctor.
-In hac habitasse platea dictumst.
-Cras ut urna dapibus elit mattis aliquam non sed augue.
-Suspendisse lacinia ornare consectetur.
-Quisque ultricies tortor non ullamcorper porta.
-Mauris ac leo quis lacus tincidunt cursus ut id felis.
+However, using external factors also poses challenges.
+Data collection for environmental parameters is resource-intensive and may require frequent updates to stay relevant.
+Additionally, interpreting these factors can be complex, as their effects vary by region and buyer priorities.
+For instance, while some buyers prioritize neighborhood aesthetics, others might value proximity to public transport more.
+Simplifying these influences into a single measure, like a geo-index, may overlook important nuances, leading to inaccuracies in valuation.
+Furthermore, reliance on advanced tools like artificial neural networks increases model complexity, requiring specialized expertise and computational power, which may not be feasible for all users.
 
@@
     "Source of the property data (e.g., MLS, Zillow).",
-    "Multiple Listing Service (MLS) identification number for the property.",
+    "Multiple Listing Service identification number for the property.",
     "Indicates whether the property is marked as a favorite (e.g., Y/N).",
@@
 
-\newpage
-
 ## Alternate Model Summaries
@@
 **TO-DO:** @tbl-models ipsum dolor sit amet, consectetur adipiscing elit.
+Duis vitae pretium velit.
+Etiam fermentum enim ac venenatis fringilla.
+Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
+Curabitur commodo pretium ornare.
+Pellentesque ultrices sapien ut lectus viverra placerat.
+Mauris nec interdum tortor.
+Cras posuere turpis sed mi blandit dictum.
+Etiam accumsan ipsum quis sem ornare cursus.
+Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
+Nullam ornare placerat risus, in luctus urna.
+Ut molestie dignissim turpis, eget tincidunt arcu.
+Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
+Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Duis vitae pretium velit.
@@
   "Reduced" = reduced_model),
+  gof_omit = "BIC|AIC|Log.Lik.",
   output = "kableExtra") |>
@@
 ```
+
+\newpage
 
@@
 
+Duis sit amet augue sagittis, commodo eros sollicitudin, congue libero.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Maecenas convallis erat velit, eu consequat massa posuere et.
+Nullam tempor purus sit amet quam lacinia euismod.
+Praesent sagittis risus sit amet odio porttitor, eget pellentesque turpis luctus.
+Proin feugiat convallis diam, id vulputate mauris.
+Nunc facilisis quam nec tristique pellentesque.
+Sed laoreet tortor at ipsum faucibus iaculis.
+Sed cursus convallis molestie.
+Aliquam et accumsan tortor.
+Morbi volutpat rutrum elit sed fermentum.
+In vestibulum placerat eros, et consectetur lacus fringilla a.
+Mauris elementum velit vel eros placerat, eget commodo nisi convallis.
+Sed ac nisi mi.
+Nullam malesuada auctor semper.
+
+Vestibulum at tortor ipsum.
+Quisque a elit ut odio suscipit fringilla ac ut nunc.
+Vivamus eget neque vel est iaculis sollicitudin in et turpis.
+Pellentesque quis risus ullamcorper metus porta viverra.
+In hac habitasse platea dictumst.
+In faucibus, sapien sed placerat semper, sapien lectus placerat leo, sit amet tempus risus dolor non erat.
+Morbi ac urna venenatis, pellentesque odio vel, iaculis tellus.
+Etiam tincidunt fringilla suscipit.
+Donec fermentum sollicitudin metus, quis tempor leo elementum eu.
+
 \newpage

### 2024-12-03 paper/paper.qmd
Updating appendix and README.md

@@
 abstract: "This paper develops a transparent property valuation model for Seattle’s real estate market, using Redfin data and statistical techniques to estimate intrinsic property values.
-The model highlights key factors influencing prices, including amenities, neighborhood characteristics, and socio-economic trends, offering an accessible tool for informed decision-making.
-Results show higher prediction errors for high-value properties and reveal time-related trends, with vintage homes gaining value due to historical significance.
-Spatial analysis captures disparities in property values, reflecting income and infrastructure differences across neighborhoods."
+Results show higher prediction errors for high-value properties and uncover time-related trends, with vintage homes gaining value due to historical significance.
+Spatial analysis captures disparities in property values, reflecting income and infrastructure differences across neighborhoods.
+The model highlights key factors influencing prices, including amenities, neighborhood characteristics, and socio-economic trends, offering a tool for informed decision-making by buyers and sellers."
 toc: true
@@
 With advancements in technology and the increasing availability of data, tools for assessing property values have become essential.
-One critical tool is an estimate of a property’s fair value, which can help buyers determine whether they are overpaying or underpaying relative to the market.
+One important tool is an estimate of a property’s fair value, which can help buyers determine whether they are overpaying or underpaying relative to the market.
 
@@
     color = "black", 
+    outlier.shape = 15, 
     width = 0.5,
-    outlier.shape = 15, 
-    outlier.size = 0.5
+    outlier.size = 0
   ) +
+  geom_point(size = 0.5, 
+             shape = 15,
+             position = position_dodge(0.6)) +
   scale_y_continuous(limits = c(0, 7)) +
@@
 Factors like historical undervaluation and socio-economic disparities remain deeply embedded, requiring more targeted interventions beyond algorithmic tools.
-Moreover, reliance on algorithmic estimates can sometimes oversimplify the nuanced factors that influence property values, such as cultural significance or community dynamics, which are more difficult to quantify.
+Moreover, reliance on algorithmic estimates can sometimes oversimplify the complex factors that influence property values, such as cultural significance or community dynamics, which are more difficult to quantify.
 Nevertheless, the use of transparent and publicly available valuation tools represents a positive step toward a more equitable real estate market.
@@
 
-Machine learning techniques have become invaluable for modeling non-linear trends in property prices, capturing the complex relationships between property characteristics, location, and market factors.
+Machine learning has become a fundamental method for modeling non-linear trends in property prices, capturing the complex relationships between property characteristics, location, and market factors.
 Traditional methods, such as linear price models, often struggle with these complexities, especially when spatial dependencies at regional levels influence property values.
@@
 
-Incorporating external environmental factors into property valuation models offers valuable advantages, such as a more comprehensive understanding of real estate prices influenced by location-specific qualities [@din] [@abelson].
+Incorporating external environmental factors into property valuation models offers great advantages, such as a more comprehensive understanding of real estate prices influenced by location-specific qualities [@din] [@abelson].
 Geographic information systems (GIS) provide detailed data on factors like neighborhood quality, proximity to amenities, and environmental conditions, enabling models to quantify their impact on property values.
@@
 For instance, while some buyers prioritize neighborhood aesthetics, others might value proximity to public transport more.
-Simplifying these influences into a single measure, like a geo-index, may overlook important nuances, leading to inaccuracies in valuation.
+Simplifying these influences into a single measure, like a geo-index, may overlook important complexities, leading to inaccuracies in valuation.
 Furthermore, reliance on advanced tools like artificial neural networks increases model complexity, requiring specialized expertise and computational power, which may not be feasible for all users.
@@
 
-**TO-DO:** @fig-diagnostics ipsum dolor sit amet, consectetur adipiscing elit.
-Duis vitae pretium velit.
-Etiam fermentum enim ac venenatis fringilla.
-Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
-Curabitur commodo pretium ornare.
-Pellentesque ultrices sapien ut lectus viverra placerat.
-Mauris nec interdum tortor.
-Cras posuere turpis sed mi blandit dictum.
-Etiam accumsan ipsum quis sem ornare cursus.
-Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
-Nullam ornare placerat risus, in luctus urna.
-Ut molestie dignissim turpis, eget tincidunt arcu.
-Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
-Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
+@fig-diagnostics presents the residual-fitted and quantile-quantile plots for the final model, used to assess the validity of linear regression assumptions.
+These diagnostics were conducted to identify any violations of key assumptions.
+Notably, two primary assumptions were tested, and one key modification was made to the model to improve its fit to the data.
+Specifically, the model was fitted using the analysis data with imputed price values to the $95^{th}$ percentile, addressing the presence of significant price outliers in the dataset.
+This adjustment is evident in the diagonal line appearing to the right in the residual-fitted plot.
+The imputation process resulted in a substantial increase in the model's $R^2$, rising from approximately 0.28 to 0.72, as detailed in @sec-model.
+The primary assumption violations observed were related to constant conditional variance and normally distributed errors.
+These issues are reflected in the left plot, where fanning indicates non-constant variance, and deviations from the diagonal line suggest departures from normality.
+To address these violations, transformations of the variables, such as using power terms, could be considered; however, this would compromise the model's interpretability.
+For the final model, interpretability was prioritized, with transformations remaining a topic for future exploration.
 
@@
 
-**TO-DO:** @tbl-models ipsum dolor sit amet, consectetur adipiscing elit.
-Duis vitae pretium velit.
-Etiam fermentum enim ac venenatis fringilla.
-Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
-Curabitur commodo pretium ornare.
-Pellentesque ultrices sapien ut lectus viverra placerat.
-Mauris nec interdum tortor.
-Cras posuere turpis sed mi blandit dictum.
-Etiam accumsan ipsum quis sem ornare cursus.
-Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
-Nullam ornare placerat risus, in luctus urna.
-Ut molestie dignissim turpis, eget tincidunt arcu.
-Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
-Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis vitae pretium velit.
-Etiam fermentum enim ac venenatis fringilla.
-Nulla imperdiet metus ac ligula ultricies, at cursus eros varius.
-Curabitur commodo pretium ornare.
-Pellentesque ultrices sapien ut lectus viverra placerat.
-Mauris nec interdum tortor.
-Cras posuere turpis sed mi blandit dictum.
-Etiam accumsan ipsum quis sem ornare cursus.
-Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus.
-Nullam ornare placerat risus, in luctus urna.
-Ut molestie dignissim turpis, eget tincidunt arcu.
-Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et.
-Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.
+@tbl-models presents the model summaries for both the full and reduced models.
+The full model includes all the variables in the analysis data, while the reduced model excludes the beds and baths variables.
+The coefficients of the two models are very similar, but the standard errors of these coefficients are slightly lower in the reduced model, suggesting lower uncertainty in the valuations.
+Regarding the goodness-of-fit metrics, the $R^2$ scores are also very similar, though the RMSE has increased.
+However, the reduced number of variables in the latter model results in adjusted $R^2$ scores that are essentially the same.
 

### 2024-12-04 paper/paper.qmd
Completing the appendix

@@
 thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
-abstract: "This paper develops a transparent property valuation model for Seattle’s real estate market, using Redfin data and statistical techniques to estimate intrinsic property values.
-Results show higher prediction errors for high-value properties and uncover time-related trends, with vintage homes gaining value due to historical significance.
-Spatial analysis captures disparities in property values, reflecting income and infrastructure differences across neighborhoods.
-The model highlights key factors influencing prices, including amenities, neighborhood characteristics, and socio-economic trends, offering a tool for informed decision-making by buyers and sellers."
+abstract: "This paper develops a transparent property valuation model for Seattle’s real estate market, using Redfin data and machine learning techniques to estimate intrinsic property values.
+Results show higher prediction errors for high-value properties and the importance of property age as well as amenities in determining valuation.
+Additionally, spatial analysis captures the influence of environmental factors on property values, reflecting income and infrastructure differences.
+These findings help buyers and sellers understand what drives property prices, supporting informed decision-making to navigate a competitive market."
 toc: true
@@
 
-**TO-DO:** Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Fusce eu hendrerit mauris, in sodales mi.
-Aenean eget urna rutrum, tristique lorem vel, maximus ipsum.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Nunc eget rutrum est.
-Donec tempor elit non nulla cursus, in consectetur est ullamcorper.
-Nam vitae porttitor felis, tempor maximus leo.
-Aenean ultricies molestie facilisis.
-Donec eget gravida eros, sit amet ornare erat.
-Cras a vehicula nibh.
-Nunc ac ligula malesuada, bibendum ante eu, luctus felis.
-Proin in tellus eu diam aliquam ultricies.
+This study uses observational data to estimate the intrinsic value of properties in Seattle based on specific parameters.
+However, observational data has limitations that can reduce valuation accuracy.
+For instance, deviations from the line of equality in @fig-scatter highlight discrepancies arising from factors not captured in the model.
+Some of these factors are difficult or impossible to measure through observational data alone, whether due to privacy concerns, challenges in data collection, or insufficient information.
+Surveys provide a valuable way to address these gaps by collecting additional data, which can help reduce the uncertainty in the model outlined in @sec-model.
+For example, surveys could capture qualitative and subjective information, such as owners’ expectations of future market trends or personal motivations for selling, that cannot be inferred from observational data.
 
-Duis sit amet augue sagittis, commodo eros sollicitudin, congue libero.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Maecenas convallis erat velit, eu consequat massa posuere et.
-Nullam tempor purus sit amet quam lacinia euismod.
-Praesent sagittis risus sit amet odio porttitor, eget pellentesque turpis luctus.
-Proin feugiat convallis diam, id vulputate mauris.
-Nunc facilisis quam nec tristique pellentesque.
-Sed laoreet tortor at ipsum faucibus iaculis.
-Sed cursus convallis molestie.
-Aliquam et accumsan tortor.
-Morbi volutpat rutrum elit sed fermentum.
-In vestibulum placerat eros, et consectetur lacus fringilla a.
-Mauris elementum velit vel eros placerat, eget commodo nisi convallis.
-Sed ac nisi mi.
-Nullam malesuada auctor semper.
+Relying solely on appraiser data presents another challenge—it often lacks the depth and variety that independent surveys can offer.
+Appraiser datasets generally focus on tangible features of the property (e.g., square footage, number of rooms) but overlook contextual or personal factors.
+Independent surveys can fill this gap by gathering details such as a seller's financial situation, their previous purchase price, or even their current perception of market conditions.
+Additionally, surveys could include questions about less commonly recorded property features like energy efficiency upgrades, aesthetic renovations, or even historical significance, which could significantly influence valuation.
+By incorporating such data into the model, a better understanding of the factors shaping a property's market value can be gained.
+This not only allows for more precise valuations but also helps assess the behavioral factors affecting seller decisions, which appraiser data alone cannot address.
 
-Vestibulum at tortor ipsum.
-Quisque a elit ut odio suscipit fringilla ac ut nunc.
-Vivamus eget neque vel est iaculis sollicitudin in et turpis.
-Pellentesque quis risus ullamcorper metus porta viverra.
-In hac habitasse platea dictumst.
-In faucibus, sapien sed placerat semper, sapien lectus placerat leo, sit amet tempus risus dolor non erat.
-Morbi ac urna venenatis, pellentesque odio vel, iaculis tellus.
-Etiam tincidunt fringilla suscipit.
-Donec fermentum sollicitudin metus, quis tempor leo elementum eu.
+Sampling is a vital component of the survey process, as it determines the reliability and representativeness of the responses.
+A well-designed sampling strategy minimizes bias and ensures diverse data collection.
+Simple random sampling is ideal in many cases because it reduces systematic bias and ensures all properties have an equal chance of inclusion.
+This method is particularly feasible given the size of the Seattle property dataset.
+However, alternative approaches like stratified sampling might be necessary if the dataset has significant variability across neighborhoods or property types.
+For instance, stratified sampling could ensure equal representation of high-value and low-value neighborhoods, or newer and older properties, which might exhibit different characteristics.
 
-Duis sit amet augue sagittis, commodo eros sollicitudin, congue libero.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Maecenas convallis erat velit, eu consequat massa posuere et.
-Nullam tempor purus sit amet quam lacinia euismod.
-Praesent sagittis risus sit amet odio porttitor, eget pellentesque turpis luctus.
-Proin feugiat convallis diam, id vulputate mauris.
-Nunc facilisis quam nec tristique pellentesque.
-Sed laoreet tortor at ipsum faucibus iaculis.
-Sed cursus convallis molestie.
-Aliquam et accumsan tortor.
-Morbi volutpat rutrum elit sed fermentum.
-In vestibulum placerat eros, et consectetur lacus fringilla a.
-Mauris elementum velit vel eros placerat, eget commodo nisi convallis.
-Sed ac nisi mi.
-Nullam malesuada auctor semper.
+Additionally, expanding the sampling pool to include properties not currently listed on the market can improve the analysis.
+Data from these properties could offer insights into external factors like neighborhood desirability, local amenities, or proximity to planned infrastructure projects, as well as internal factors like deferred maintenance or owner sentiment about selling.
+A multi-stage sampling process could be employed, where neighborhoods are first selected based on certain characteristics (e.g., median income levels or recent development activity), followed by a random selection of properties within those neighborhoods.
+Responses from this expanded sample would help tune the model to reflect a broader range of scenarios, ultimately improving its robustness and applicability.
 
-Vestibulum at tortor ipsum.
-Quisque a elit ut odio suscipit fringilla ac ut nunc.
-Vivamus eget neque vel est iaculis sollicitudin in et turpis.
-Pellentesque quis risus ullamcorper metus porta viverra.
-In hac habitasse platea dictumst.
-In faucibus, sapien sed placerat semper, sapien lectus placerat leo, sit amet tempus risus dolor non erat.
-Morbi ac urna venenatis, pellentesque odio vel, iaculis tellus.
-Etiam tincidunt fringilla suscipit.
-Donec fermentum sollicitudin metus, quis tempor leo elementum eu.
+The survey will be administered in person, where surveyors will knock on each household in the sample's door to ensure detailed and accurate data collection, allowing interviewers to clarify questions.
+Participants will include property owners, both those currently selling and those who are not, to gather diverse perspectives.
+The survey will feature questions such as:
+
+- "What year did you purchase the property, and at what price?"
+- "Have there been any significant renovations or upgrades?"
+- "Is the property fully furnished?"
+- "What motivated you to sell (if applicable)?"
+- "How would you rate your neighborhood on factors like safety, accessibility, and amenities?"
+- "What price range do you believe is fair for this property?"
+
+These questions aim to uncover financial, sentimental, and qualitative factors influencing property valuation.
+This survey shares similarities with the American Housing Survey, which is another household survey that interviews all types of residents such as property managers, regular occupants, etc.
+[@ahs]
+
+Challenges to measurement include potential biases, such as response bias, where respondents may provide answers they perceive as favorable rather than accurate (e.g., overstating the quality of renovations).
+Additionally, respondents may have incomplete or inaccurate records of past transactions or upgrades.
+Interviewer bias is another concern, as the phrasing or tone of questions might unintentionally influence responses.
+Addressing these challenges involves careful question design, interviewer training, and cross-referencing self-reported data with available public records where possible.
 

### 2024-12-04 scripts/cleaning.R
Adding raw data

@@ -24,6 +24,9 @@ raw_5bed <- read_csv("data/raw_data/raw_5bed.csv")
 # Merging the data
 raw_data <- rbind(raw_0bed, raw_1bed, raw_2bed, raw_3bed, raw_4bed, raw_5bed)
 
+# Exporting raw data
+write_csv(raw_data, "data/raw_data/raw_data.csv")
+
 # Cleaning the data
 cleaned_data <- raw_data |>
   clean_names() |>

### 2024-12-04 other/datasheet/datasheet.qmd
Completing the datasheet

---
title: "Datasheet for 'raw_data.csv'"
author: 
  - Sameeck Bhatia
date: today
date-format: long
thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
abstract: ""
format: pdf
number-sections: true  
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was created to allow real estate agents and the public to view current property listings in the city of Seattle.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   The dataset was created by the real estate agents of Seattle on behalf of their brokerage firms, who, in turn, created it for the National Association of Realtors.
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   The creation of the dataset has been funded by real estate brokerage firms who pay for appraisal services.
4.  *Any other comments?*
    -   The observations in the dataset were collected by appraisers and reported to real estate agents and brokers, who have then entered this data into the Multiple Listing Service.

**Composition**

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   Each instance represents a unique property that has been listed for sale.
2.  *How many instances are there in total (of each type, if appropriate)?*
    -   There are 1403 instances/listings.
3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   The dataset is a sample of a larger set. The population is all current property listings of all types (e.g. townhouse, land, multi-family, etc.) in Seattle.
4.  *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
    -   Each instance consists of features that define the property. There features range from the number of days the property has been up on the market to geographic coordinates.
5.  *Is there a label or target associated with each instance? If so, please provide a description.*
    -   There is a label associated for every instance. The labels (variables in the data) have been outlined in `paper.qmd` in the appendix.
6.  *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
    -   There are some instances missing that are labeled by `ADDRESS`, `LOCATION`, `SQUARE FEET`, `YEAR BUILT`, `HOA/MONTH`. The instances in `ADDRESS` and `LOCATION` are related and are missing due to the address of the listings being kept private. One instance of `SQUARE FEET` is missing and is likely due to no appraiser measuring the size yet. A few instances in `YEAR BUILT` are missing as the information has likely not been recorded. Lastly, many instances of `HOA/MONTH` are missing as most single-family homes do not require homeowner's association fees.
7.  *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
    -   There are no explicit relationships between individual instances.
8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   One recommended split it having training data cover the outliers in price so that a model trained on it does not underfit the testing data. There is no specific proportion for a train-test split.
9.  *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
    -   There are no errors and no duplicates in the dataset.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
    -   The dataset is not self-contained. All instances have a link to the individual listing's website. There is a guarantee that the link will exist in the future as the data source (Redfin) ensures the listing is updating to reflect its status (e.g. active, sold, sold conditional). There are no official archives of the dataset.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
    -   There is no data that is confidential as there is no personal identifiable information published onto the Multiple Listing Service.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
    -   The dataset contains no instances that might be offensive of threatening.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
    -   The dataset does not contain data on a sub-population.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
    -   It is not possible to identify individuals with any instance in the dataset as there is no PII (Personal Identifiable Information).
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
    -   The dataset contains no instances that are sensitive.
16. *Any other comments?*
    -   None.

**Collection process**

1.  *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
    -   The data was directly observable by the appraiser who reported each instance.
2.  *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
    -   The data was manually collected by an appraiser, a human that measures each feature of the data.
3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   The dataset is a sample of a larger set and it was collected using non-probabilistic sampling. The dataset was sampled by the most recent listings on the MLS.
4.  *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
    -   Appraisers, real estate agents/brokers, and the seller of each property was involved. The appraiser was employed by either the seller of the broker and collected the data with consent of the seller. There is no information on the exact amount each appraiser was paid, however, in Seattle they are typically paid \$600 - \$800 per appraisal.
5.  *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
    -   The timeframe for data collection is unknown as the appraisal could have been conducted on any day and the information could have been entered in the MLS database on another.
6.  *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   The American Society of Appraisers usually conducts an ethical review for the data collection methods and reviews each appraiser's compliance with the regulations.
7.  *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
    -   The data was collected via a third party that collects it from the Multiple Listing Service database (Redfin).
8.  *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
    -   There are no individuals in question for the dataset.
9.  *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
    -   There are no individuals in question for the dataset.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
    -   There are no individuals in question for the dataset.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   Analysis on this particular dataset and its impact has not been conducted as there are no particular subjects in question.
12. *Any other comments?*
    -   None.

**Preprocessing/cleaning/labeling**

1.  *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
    -   There was no preprocessing done on the data as it is a raw dataset.
2.  *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
    -   Not applicable.
3.  *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
    -   Not applicable.
4.  *Any other comments?*
    -   None.

**Uses**

1.  *Has the dataset been used for any tasks already? If so, please provide a description.*
    -   This particular dataset has not been used for any tasks. However, similar versions of the dataset (from Realtor.com and Zillow) has been used for time-series price prediction analysis.
2.  *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
    -   There is no one repository that links to the system that uses the dataset as it is consumed by multiple users.
3.  *What (other) tasks could the dataset be used for?*
    -   The dataset can be used to identify areas where a potential buyer would want to purchase their property.
4.  *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
    -   The dataset contains some information regarding the terms of use that is not directly relevant to the instance of the data for analysis.
5.  *Are there tasks for which the dataset should not be used? If so, please provide a description.*
    -   The dataset should not be used for advertising campaigns and non real-estate activities such as reposting active listings and overloading the MLS system.
6.  *Any other comments?*
    -   None.

**Distribution**

1.  *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
    -   The dataset will be distributed to real estate management and brokerage companies as a form of collaboration.
2.  *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
    -   There is no information on how the dataset will be distributed.
3.  *When will the dataset be distributed?*
    -   There is no information on when the dataset will be distributed.
4.  *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
    -   There is no information on the copyright that the dataset will be distributed under.
5.  *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
    -   There are no IP-based restrictions on the data.
6.  *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
    -   There is one export control. The control is the condition that the individual extracting the data has a Redfin account.
7.  *Any other comments?*
    -   None.

**Maintenance**

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   The dataset will be mainted by Redfin as the company is constantly extracting data from the MLS database.
2.  *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
    -   The owner of the dataset cannot be contacted as the contact information is not publicly available.
3.  *Is there an erratum? If so, please provide a link or other access point.*
    -   There is no erratum.
4.  *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
    -   The dataset will be updated. It is updated every day on the Redfin website.
5.  *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
    -   The dataset does not relate to people.
6.  *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
    -   Older versions of the dataset will be kept in a separate database and can be accessed by request from the source (Redfin).
7.  *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
    -   There is definitely a way to build on the dataset. The dataset has been collected at a given point in time (cross-sectional data) and any data collected after that point is a natural addition to the dataset.
8.  *Any other comments?*
    -   None.

\newpage

# References
